---
title: "Canola Bee Analysis"
author: "Dr. Riley M. Anderson, Olivia Shaffer, & Salena Helmreich"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

This analysis explores Salena and Olivia's canola experiment.

### Summary of Results

* No difference in bee community composition across habitat types (natural or semi-natural), or habitat types and bloom period.

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse)
library(cowplot) 
library(glmmTMB)
library(lme4)
library(sjPlot)
library(knitr)
library(emmeans)
library(vegan)
library(randomForest)
library(caret)
library(adespatial)
library(ggrepel)

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
bees <- read.csv("data/Bee_Samples_2022.csv")

plants <- read.csv("data/PlantDiv2022.csv")

bee_hw <- read.csv("data/CleanPalouseBees.csv")

canola <- read.csv("data/canola_coverage.csv")

canola_bees <- read.csv("data/bees_in_canola.csv")

###### Functions


```


```{r Data_Wrangling, echo = F, comment = ""}

# clean up Salena's data:
bee_hw <- bee_hw %>% 
  mutate(across(
    c(period : land_type), factor),
    across(c(transect : genus), factor),
    date = ymd(date),
    year = as.numeric(year(date)) + 2000,
    month = as.numeric(month(date)),
    day = as.numeric(day(date)),
    date = paste(year, month, day, sep = "_"),
    date = ymd(date),
    site = factor(case_when(
      site == "Smoot " ~ "Smoot",
      TRUE ~ site)),
    period = factor(case_when(
      period == "1" ~ "Pre-Bloom",
      period == "2" ~ "Peak-Bloom",
      TRUE ~ "Post-Bloom")),
    in_canola = factor("no"))


hw_1 <- bee_hw %>% 
  select(date, period, site, land_type, latitude, longitude,
         transect, morphogroup, family, genus, head_width,
         intertegular_space, in_canola)

canola_bees <- canola_bees %>% 
  filter(!is.na(ID),
         Genus != "Wasp",
           Genus != "") %>% 
  mutate(across(c(
    Site : sex, County : Catch), factor)) %>% 
  unite(new_date, c(Year, Month, Day), sep = "-") %>% 
  mutate(date = ymd(new_date),
         Period = factor(case_when(
      Period == "2" ~ "Peak-Bloom",
      Period == "3" ~ "Post-Bloom",
      TRUE ~ NA)),
      land_type = factor("canola"),
      in_canola = factor("yes"),
      HW = as.numeric(HW)) %>% 
  select(-Date) %>% 
  rename(site = Site,
         period = Period,
         transect = Transect,
         genus = Genus,
         species = Species,
         head_width = HW,
         latitude = Lat,
         longitude = Long,
         subgenus = Subgenus)

hw_2 <- canola_bees %>% 
  select(date, period, site, land_type, latitude, longitude,
         transect, genus, subgenus, species, head_width,
         land_type, in_canola)

bee_head_width <- bind_rows(hw_1, hw_2)


#{{{{{{{{{{{{}}}}}}}}}}}}
# plant data:
plants <- plants %>% 
  filter(Genus != "UK") %>% 
  mutate(Date = mdy(Date),
         across(c(Period : Species), factor),
         Period = factor(case_when(
           Period == "PreBloom" ~ "Pre-Bloom",
           Period == "PeakBloom" ~ "Peak-Bloom",
           Period == "DeclineBloom" ~ "Post-Bloom",
           TRUE ~ NA)),
         Site = factor(case_when(
           Site == "TurnbullPlaza" ~ "Plaza",
           Site == "TurnbullSpangle" ~ "Spangle",
           TRUE ~ Site)),
         siteID = factor(paste(Site, Period, sep = "_")),
         FloralUnits = case_when(
           is.na(FloralUnits) ~ 0,
           TRUE ~ FloralUnits),
         DoY = yday(Date),
         DoY = case_when(
           DoY == 153 & Site == "Chipman1" ~ 151,
           DoY == 153 & Site == "Chipman2" ~ 151,
           TRUE ~ DoY)) %>% 
  unite(spp, c(Genus, Species), sep = " ", remove = F) %>% 
  rename(date = Date,
         period = Period,
         site = Site,
         transect = Transect,
         family = Family,
         genus = Genus,
         species = Species)

canola_plants <- data.frame(
  site = factor(levels(canola_bees$site)),
  CommonName = factor("canola"),
  Amount = 100,
  FloralUnits = 100)


plants <- bind_rows(plants, canola_plants)

#########################

# Site data
canola <- canola %>% 
  mutate(site = factor(site),
         PrairieFragment = factor(PrairieFragment),
         PrairieFragment = factor(case_when(
           site == "Asotin" | site == "Hutchins" ~ "No",
           site == "Plaza" | site == "Spangle" ~ "Yes",
           TRUE ~ PrairieFragment
         )))

canola_sites <- bee_head_width %>%
  filter(in_canola == "yes") %>%
  select(site, latitude, longitude) %>%
  distinct() %>%
  mutate(lat_long = paste(latitude,longitude, sep = ",")) %>% 
  filter(!is.na(latitude))

write.csv(canola_sites,
          file = "data/canola_site_coords.csv",
          quote = F, row.names = F)





########## Probably discard this below:
plant_counts <- plants %>% 
  uncount(Amount) %>% 
  group_by(site, period, CommonName) %>% 
  tally(name = "plant_count")


canola <- canola %>% 
  mutate(site = factor(site))

canola_vars <- colnames(canola)

canola_bee_vars <- data.frame(levels(canola_bees$site), 1:1, 0:0, 0:0, 0:0,
                              c(NA,NA,NA,NA,NA,NA,NA,NA,NA))

colnames(canola_bee_vars) <- canola_vars

canola_bee_vars <- canola_bee_vars %>% 
  mutate(site = factor(site))

canola <- canola %>% 
  bind_rows(., canola_bee_vars)


year_day <- plants %>% 
  select(siteID, DoY) %>% 
  distinct()

#######################################################

# Which genera are found in both natural and semi-natural sites?

# nat <- bees1 %>% 
#   filter(land_type == "natural") %>% 
#   select(genus) %>% 
#   distinct()
# 
# seminat <- bees1 %>% 
#   filter(land_type == "semi") %>% 
#   select(genus) %>% 
#   distinct()
# 
# # species only found in NATURAL sites:
# bees_unique_natural <- anti_join(nat, seminat)
# 
# # species only found in SEMINATURAL sites:
# bees_unique_semi <- anti_join(seminat, nat)
# 
# # overlapping species (species we can use in site comparison models)
# overlap <- semi_join(nat, seminat, by = "genus")
# 
# # removing species that were only found in either natural or seminatural sites:
# bees.reduced <- semi_join(bees1, overlap, by = "genus")
# 
# 
# # removing genera with sample size < 3 in either land type:
# good_samples <- bees.reduced %>% 
#   group_by(land_type, genus) %>% 
#   tally() %>% 
#   filter(n > 2) %>% 
#   ungroup() %>% 
#   select(genus) %>% 
#   distinct()
# 
# 
# bees.reduced <- semi_join(bees.reduced, good_samples, by = "genus")



```


```{r plant_community_wrangling, echo = F}

plant_matrix <- plants %>% 
  uncount(Amount) %>% 
  select(site, CommonName) %>% 
  group_by(site, CommonName) %>% 
  tally() %>% 
  pivot_wider(names_from = CommonName,
              values_from = n,
              values_fill = list(n = 0)) %>% 
  column_to_rownames("site")

plant_div <- cbind(
  shannon = diversity(plant_matrix),
  richness = specnumber(plant_matrix),
  pielou = diversity(plant_matrix) / log(specnumber(plant_matrix))
) %>% as.data.frame() %>% 
  rownames_to_column("site") %>% 
  mutate(site = factor(site))

total_floral <- plants %>%
  group_by(site) %>% 
  summarise(cu_floral = sum(FloralUnits))

plant_div <- left_join(plant_div, total_floral,
                       by = "site")
  

```

## Plant community composition

### Plant diversity across bloom period
```{r plant_diversity_bloomperiod, echo = F}
plant_div_fun <- function(data, bloom_period){
  data %>% 
    uncount(Amount) %>% 
    filter(period == bloom_period) %>% 
    group_by(site, spp) %>% 
    tally(name = "plant_count") %>% 
    pivot_wider(names_from = spp,
              values_from = plant_count,
              values_fill = list(plant_count = 0)) %>% 
    column_to_rownames("site") %>% 
    diversity()
}

period_plant_div <- levels(plants$period) %>% 
  map(~ plant_div_fun(plants, bloom_period = .x))

names(period_plant_div) <- levels(plants$period)

plant_div <- do.call(rbind, period_plant_div) %>% 
  data.frame() %>% 
  rownames_to_column("period") %>% 
  pivot_longer(cols = c("Asotin" : "Spangle"),
               names_to = "site",
               values_to = "div") %>% 
  mutate(across(c(period, site), factor),
         period = factor(period, levels = c(
           "Pre-Bloom", "Peak-Bloom", "Post-Bloom"
         ))) %>% 
  left_join(., land_type, by = "site") %>% 
  left_join(., canola, by = "site") %>% 
  mutate(siteID = factor(paste(site, period, sep = "_"))) %>% 
  left_join(., year_day, by = "siteID")

plant_div %>% 
  ggplot(aes(x = period, y = div)) +
  geom_boxplot() +
  theme_cowplot(font_size = 16) +
  labs(y = "Shannon diversity") +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~land_type)

div_model1 <- lm(div ~ period * land_type,
                 data = plant_div)

anova(div_model1)
summary(div_model1)
```


```{r plant_div_over_time, echo = F}

plant_div %>% 
  ggplot(aes(x = DoY, y = div, color = land_type,
             fill = land_type)) +
  geom_jitter(width = 1, height = 0.2, shape = 21,
              aes(fill = land_type), color = 'black', size = 3) +
  geom_smooth(method = 'lm', alpha = 0.2) +
  theme_classic()

div_model2 <- lm(div ~ DoY * land_type,
                 data = plant_div)

anova(div_model2)
summary(div_model2)
```

# Multivariate analysis of plant species turnover across bloom period and habitat types

```{r multi_plants, echo = F, fig.width=10, fig.height=8}

plant_totals <- plants %>% 
  uncount(Amount) %>% 
  group_by(siteID, spp) %>% 
  tally() %>% 
  group_by(spp) %>% 
  summarise(total_counts = sum(n)) %>% 
  arrange(desc(total_counts)) %>% 
  filter(total_counts > quantile(total_counts, 0.1))


meta_plants <- plants %>% 
  select(siteID, period, site, land_type, DoY) %>% 
  distinct()


plant_matrix <- plants %>% 
  semi_join(., plant_totals, by = "spp") %>% 
  uncount(Amount) %>% 
  group_by(siteID, spp) %>% 
  tally() %>% 
  pivot_wider(names_from = spp,
              values_from = n,
              values_fill = list(n = 0)) %>% 
  column_to_rownames("siteID")

plant_perm <- adonis2(plant_matrix ~ period * land_type,
                      data = meta_plants)

plant_perm

simper_plants <- simper(plant_matrix,
                        group = meta_plants$land_type)

summary(simper_plants)

plant_lt_cap <- capscale(plant_matrix ~ land_type,
                         data = meta_plants,
                         distance = "bray")
summary(plant_lt_cap)


site_scores <- scores(plant_lt_cap, display = "sites") %>%
  as.data.frame() %>%
  mutate(sample = rownames(.),
         land_type = meta_plants$land_type)

species_scores <- scores(plant_lt_cap, display = "species") %>%
  as.data.frame() %>%
  rownames_to_column("species")

# Optional: keep only species with strongest loadings on CAP1
species_scores_filtered <- species_scores %>%
  filter(abs(CAP1) > 0.1 | abs(MDS1) > 0.1)

ggplot(site_scores,
       aes(x = CAP1, y = MDS1, color = land_type, fill = land_type)) +
  stat_ellipse(geom = "polygon", type = "t",
               level = 0.95, alpha = 0.3) +
  geom_point(size = 2.5, shape = 21, stroke = 0.5,
             aes(fill = land_type), color = 'black') +
  geom_segment(data = species_scores_filtered,
               aes(x = 0, y = 0, xend = CAP1, yend = MDS1),
               arrow = arrow(length = unit(0.2, "cm")),
               color = "black",
               inherit.aes = F) +
  geom_text_repel(data = species_scores_filtered,
            aes(x = CAP1 * 1.1, y = MDS1 * 1.1, label = species),
            size = 3, hjust = 0.5, max.overlaps = 20,
            inherit.aes = F) +
  scale_color_manual(values = c("natural" = "#1b9e77",
                                "semi" = "#d95f02"),
                     labels = c("Natural", "Semi-natural")) +
  scale_fill_manual(values = c("natural" = "#1b9e77",
                                "semi" = "#d95f02"),
                     labels = c("Natural", "Semi-natural")) +
  theme_minimal(base_size = 14) +
  labs(
    x = "CAP1",
    y = "MDS1",
    fill = "Land type", color = "Land type")

```

* No differences in plant community across bloom period or land type.

This may be expected, as the plants don't get up and walk away...certainly not within a season anyway.

**However, we should expect floral resources to vary across bloom periods.**

### Plant floral diversity across bloom period
```{r floral_diversity_bloomperiod, echo = F}
floral_div_fun <- function(data, bloom_period){
  data %>% 
    uncount(FloralUnits) %>% 
    filter(period == bloom_period) %>% 
    group_by(site, spp) %>% 
    tally(name = "plant_count") %>% 
    pivot_wider(names_from = spp,
              values_from = plant_count,
              values_fill = list(plant_count = 0)) %>% 
    column_to_rownames("site") %>% 
    diversity()
}

period_floral_div <- levels(plants$period) %>% 
  map(~ floral_div_fun(plants, bloom_period = .x))

names(period_floral_div) <- levels(plants$period)

floral_div <- do.call(rbind, period_floral_div) %>% 
  data.frame() %>% 
  rownames_to_column("period") %>% 
  pivot_longer(cols = c("Asotin" : "Spangle"),
               names_to = "site",
               values_to = "div") %>% 
  mutate(across(c(period, site), factor),
         period = factor(period, levels = c(
           "Pre-Bloom", "Peak-Bloom", "Post-Bloom"
         ))) %>% 
  left_join(., land_type, by = "site") %>% 
  left_join(., canola, by = "site")

floral_div %>% 
  ggplot(aes(x = period, y = div)) +
  geom_boxplot() +
  theme_cowplot(font_size = 16) +
  labs(y = "Floral Shannon diversity") +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~land_type)

floral_model1 <- lm(div ~ period * land_type,
                 data = floral_div)

anova(floral_model1)
summary(floral_model1)

```

```{r floral_units_over_time, echo = F}

total_floral <- plants %>% 
  group_by(siteID, DoY, land_type) %>% 
  summarise(total_FU = sum(FloralUnits)) 

total_floral %>% 
  ggplot(aes(x = DoY, y = total_FU, color = land_type)) +
  geom_jitter(width = 0.1, height = 0) +
  geom_smooth(method = 'lm') +
  theme_classic() +
  labs(x = "Day of year", y = "Total floral units",
       color = "Land type")

floral_mod_total <- lm(total_FU ~ DoY * land_type,
                       data = total_floral)

anova(floral_mod_total)
summary(floral_mod_total)
```

# Multivariate analysis of floral resources across bloom periods

```{r multi_floral, echo = F}

floral_matrix <- plants %>% 
  uncount(FloralUnits) %>% 
  group_by(siteID, spp) %>% 
  tally() %>% 
  pivot_wider(names_from = spp,
              values_from = n,
              values_fill = list(n = 0)) %>% 
  column_to_rownames("siteID")

floral_perm <- adonis2(floral_matrix ~ period * land_type,
                      data = meta_plants)

floral_perm

```



### Plant community diversity across bloom period and canola
```{r plant_diversity_bloom_canola, echo = F}

plot_canola_radius <- plant_div %>% 
  ggplot(aes(x = propCan1km, y = div)) +
  geom_jitter(width = 0.01, shape = 21, size = 3, stroke = 0.7,
             aes(fill = land_type)) +
  labs(y = "Shannon diversity", x = "Pr(Canola) 1km radius",
       fill = "Habitat type") +
  theme_bw() +
  theme(strip.background = element_blank()) +
  facet_wrap(~period)


plot_canola_distance <- plant_div %>% 
  ggplot(aes(x = DistanceToCanola, y = div)) +
  geom_jitter(width = 0.01, shape = 21, size = 3, stroke = 0.7,
             aes(fill = land_type)) +
  labs(y = "Shannon diversity", x = "Distance to canola (m)",
       fill = "Habitat type") +
  theme_bw() +
  theme(strip.background = element_blank()) +
  facet_wrap(~period) +
  scale_x_continuous(transform = "log10")

cowplot::plot_grid(plot_canola_radius,
          plot_canola_distance,
          nrow = 2)

```



### Turnover of floral resources

$\beta$ diversity

```{r plant_beta_div_across_bloom_period, echo = F}
plant_count_matrix <- plant_counts %>% 
  mutate(site_period = factor(paste0(site, period, sep = "_"))) %>% 
  ungroup() %>% 
  pivot_wider(names_from = CommonName,
              values_from = plant_count,
              values_fill = list(plant_count = 0))


fun_beta_div <- function(data, site_name){
beta_div_list <- data %>% 
  filter(site == site_name) %>% 
  select(-site, -site_period) %>% 
  column_to_rownames("period") %>% 
  beta.div.comp(., coef = "S", quant = TRUE)

beta_div_parts <- beta_div_list["part"] %>% data.frame %>% t() %>%
  data.frame() 

rownames(beta_div_parts) <- site_name

beta_div_parts <- beta_div_parts %>% rownames_to_column("site")
}

plant_beta_div <- levels(plant_count_matrix$site) %>% 
  map(~ fun_beta_div(plant_count_matrix, site_name = .x))

plant_beta_div <- do.call(rbind, plant_beta_div) %>% 
  left_join(., land_type, by = "site") %>% 
  left_join(., canola, by = "site")

```

```{r plant_betadiv_by_site_fig, echo = F}

plant_beta_div %>% 
  pivot_longer(cols = c(BDtotal:RichDif.BDtotal),
               names_to = "beta_types",
               values_to = "beta_parts") %>% 
  mutate(site = factor(site, levels = c(
    "Chipman1", "Chipman2", "Clark1", "Hutchins",
    "Asotin", "Clark2", "Kamiak", "Plaza", "Smoot", "Spangle"))) %>% 
  ggplot(aes(x = site, y = beta_parts, color = land_type)) +
  geom_point() +
  facet_wrap(~ beta_types) +
  coord_flip() +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank()) +
  labs(color = "Habitat")
```

```{r plant_betadiv_prop_canola_fig, echo = F}


plant_beta_div %>% 
  pivot_longer(cols = c(BDtotal:RichDif.BDtotal),
               names_to = "beta_types",
               values_to = "beta_parts") %>% 
  mutate(site = factor(site, levels = c(
    "Chipman1", "Chipman2", "Clark1", "Hutchins",
    "Asotin", "Clark2", "Kamiak", "Plaza", "Smoot", "Spangle"))) %>% 
  ggplot(aes(x = propCan1km, y = beta_parts, color = land_type)) +
  geom_point() +
  facet_wrap(~ beta_types) +
  theme_bw() +
  theme(strip.background = element_blank()) +
  labs(color = "Habitat",
       y = "Plant species turnover")
```

## Bee diversity across bloom periods
```{r bee_div_bloom_period, echo = F}


bee_div_fun <- function(data, bloom_period){
  data %>% 
    filter(period == bloom_period) %>% 
    group_by(site, genus) %>% 
    tally(name = "bee_count") %>% 
    pivot_wider(names_from = genus,
              values_from = bee_count,
              values_fill = list(bee_count = 0)) %>% 
    column_to_rownames("site") %>% 
    diversity()
}

period_bee_div <- levels(bee_div_data$period) %>% 
  map(~ bee_div_fun(bee_div_data, bloom_period = .x))

names(period_bee_div) <- levels(bees$period)

bee_div <- do.call(rbind, period_bee_div) %>% 
  data.frame() %>% 
  rownames_to_column("period") %>% 
  pivot_longer(cols = c("Asotin" : "Smoot_c"),
               names_to = "site",
               values_to = "div") %>% 
  mutate(across(c(period, site), factor),
         period = factor(period, levels = c(
           "Pre-Bloom", "Peak-Bloom", "Post-Bloom"
         ))) %>% 
  left_join(., land_type, by = "site") %>% 
  left_join(., canola, by = "site")

```

## Bee turnover across bloom periods
```{r bee_beta_div_period, echo = F}

fun_beta_div <- function(data, site_name){
beta_div_list <- data %>% 
  filter(site == site_name) %>% 
  select(-site, -site_period) %>% 
  column_to_rownames("period") %>% 
  beta.div.comp(., coef = "S", quant = TRUE)

beta_div_parts <- beta_div_list["part"] %>% data.frame %>% t() %>%
  data.frame() 

rownames(beta_div_parts) <- site_name

beta_div_parts <- beta_div_parts %>% rownames_to_column("site")
}

bee_beta_div <- levels(bee_count_matrix$site) %>% 
  map(~ fun_beta_div(bee_count_matrix, site_name = .x))

bee_beta_div <- do.call(rbind, bee_beta_div) %>% 
  left_join(., land_type, by = "site") %>% 
  left_join(., canola, by = "site")


```

# Testing Olivia's predictions

1) The stability of bee communities (measured as turnover from the pre- to peak- to post-bloom) will be greater in landscapes with less access to pulse-blooming resources. **Evidence in support of this prediction will be a positive relationship between $\beta$ diversity and the proportion of canola surrounding each site or a negative relationship between $\beta$ diversity and the distance to the nearest canola.**


```{r prediction_1, echo = F}

bee_beta_div %>% 
  pivot_longer(cols = c(BDtotal : RichDif.BDtotal),
               names_to = "beta_types",
               values_to = "beta_div") %>% 
  ggplot(aes(x = propCan1km, y = beta_div)) +
  geom_point() +
  geom_smooth(alpha = 0.2,
              method = lm, formula = y ~ x + poly(x, 2)) +
  theme_bw() + facet_wrap(~ beta_types)

bee_beta_div %>% 
  pivot_longer(cols = c(BDtotal : RichDif.BDtotal),
               names_to = "beta_types",
               values_to = "beta_div") %>% 
  ggplot(aes(x = propCan500m, y = beta_div)) +
  geom_point() +
  geom_smooth(alpha = 0.2,
              method = lm, formula = y ~ x + poly(x, 2)) +
  theme_bw() + facet_wrap(~ beta_types)

bee_beta_div %>% 
  pivot_longer(cols = c(BDtotal : RichDif.BDtotal),
               names_to = "beta_types",
               values_to = "beta_div") %>% 
  ggplot(aes(x = propCan2km, y = beta_div)) +
  geom_point() +
  geom_smooth(alpha = 0.2,
              method = lm, formula = y ~ x + poly(x, 2)) +
  theme_bw() + facet_wrap(~ beta_types)

bee_beta_div %>% 
  pivot_longer(cols = c(BDtotal : RichDif.BDtotal),
               names_to = "beta_types",
               values_to = "beta_div") %>% 
  ggplot(aes(x = DistanceToCanola/1000, y = beta_div)) +
  geom_point() +
  geom_smooth(alpha = 0.2,
              method = lm, formula = y ~ x + poly(x, 2)) +
  theme_bw() + facet_wrap(~ beta_types) +
  labs(x = "Distance to Canola (km)")
```

(2) Bee communities at sites with more canola in the landscape will have less stability due to a reliance on pulsed floral resources 

**How is this different from the first prediction?**


(3) Bee communities will be more diverse during peak bloom than pre- and post-bloom
```{r bee_divesity_over_bloom_period, echo = F}

bee_div <- bee_div %>% 
  mutate(land_type = factor(case_when(
    land_type == "natural" ~ "natural",
    land_type == "semi" ~ "semi",
    TRUE ~ "canola"
  )))

bee_div %>% 
  ggplot(aes(x = period, y = div)) +
  geom_boxplot() +
  geom_jitter(width = 0.1) +
  theme_bw() + labs(y = "Bee diversity") +
  facet_wrap(~ land_type)

bee_div_mod1 <- lm(div ~ period * land_type, data = bee_div)

summary(bee_div_mod1)

plot_model(bee_div_mod1,
           type = "pred",
           terms = c("land_type", "period"))
```

4) Bee health metrics will be greater in landscapes with more canola
```{r bee_health_by_canola, echo = F}


head_width <- bee_hw %>% 
  select(date, site, period, genus, head_width, land_type)

can_head_width <- canola_bees %>% 
  select(date, site, period, genus, head_width, land_type) %>% 
  filter(!is.na(head_width)) %>% 
  mutate(head_width = as.numeric(head_width))

head_width <- head_width %>% 
  bind_rows(., can_head_width)


head_width <- head_width %>% 
  left_join(., canola, by = "site")

head_width_mod1 <- lmer(head_width ~ propCan500m +
                          scale(DistanceToCanola) +
                          (propCan500m | genus), data = bee_hw)
summary(head_width_mod1)
plot_model(head_width_mod1,
           type = "re") +
  geom_vline(xintercept = 0, linewidth = 1, color = "black")


plot_model(head_width_mod1, type = "diag")
```

(5) Bee health metrics will be positively related to stability of floral resources in the landscape

* can't make this comparison without plant diversity data in canola fields. Or we make an assumption about low diversity (1 plant, low richness, low evenness)
```{r bee_health_by_plant_betadiv, echo = F}




```


## Bee community composition

```{r bee_composition_by_canola, echo = F}

meta_plants <- meta_plants %>% 
  left_join(., canola, by = "site")

# PERMANOVA be composition by predictors

## need to combine canola sites with plant meta data for ordinations

# bee_comp_perm <- adonis2(bee_count_matrix ~ )

bee_ord_model1 # Only prop canola matters

## checking for trends with GAM:
library(mgcv)
bee_mat_gam <- bind_cols(bee_mat, scores(bee_ord)$sites)

bee_gam1 <- gam(log(DistanceToCanola) ~ s(NMDS1, NMDS2, NMDS3, k = 9) +
                  period,
                  data = bee_mat_gam)

summary(bee_gam1) # Distance to canola has strong relationship

bee_gam2 <- gam(propCan1km ~ s(NMDS1, NMDS2, NMDS3, k = 2) +
                  period,
                  data = bee_mat_gam)

summary(bee_gam2) # prop Canola has slight relationship


#### checking for trends with PERMANOVA:

bee_mat_species <- bee_mat %>% ungroup() %>% 
  select(Andrena : Anthophora)

bee_mat_predictors <- bee_mat %>% ungroup() %>% 
  select(-Andrena : -Anthophora)

# linear fit:
bee_ord_adonis1 <- adonis2(bee_mat_species ~
                            log(DistanceToCanola) +
                            propCan1km + period,
                          data = bee_mat)


# checking for non-linearity in DistanceToCanola
bee_ord_adonis2 <- adonis2(bee_mat_species ~
                            splines::ns(log(DistanceToCanola), df = 3) +
                            propCan1km + period,
                          data = bee_mat)

# checking for non-linearity in propCan1km
bee_ord_adonis3 <- adonis2(bee_mat_species ~
                            splines::ns(propCan1km, df = 3) +
                            DistanceToCanola + period,
                          data = bee_mat)

# checking for non-linearity in propCan1km & DistanceToCanola
bee_ord_adonis4 <- adonis2(bee_mat_species ~
                            splines::ns(propCan1km, df = 3) +
                            splines::ns(DistanceToCanola, df = 3) +
                             period,
                          data = bee_mat)
bee_ord_adonis1
bee_ord_adonis2
bee_ord_adonis3
bee_ord_adonis4

# Both propCan1km and DistanceToCanola have non-linear fits
```

```{r bee_comp_by_canola_plot, echo = F, fig.width=9, fig.height=10}

bee_sites <- data.frame(scores(bee_ord)$sites) %>% 
  rownames_to_column("sample_num") %>% 
  bind_cols(., bee_mat_predictors) %>% 
  mutate(period = factor(period, levels = c(
    "Pre-Bloom", "Peak-Bloom", "Post-Bloom"
  )))

bee_sites_centroids <- bee_sites %>% 
  group_by(period) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2)) 

bee_ord_propCan <- gam(propCan1km ~ s(NMDS1, NMDS2, bs = "ts"),
                       data = bee_sites)
summary(bee_ord_propCan)
bee_ord_distance <- gam(log(DistanceToCanola) ~ s(NMDS1, NMDS2,
                                                  bs = "ts"),
                        data = bee_sites)

preddat <- expand.grid(NMDS1 = seq(-2.5, 2.5, 0.5),
                       NMDS2 = seq(-1.5, 1.5, 0.5))
preddat$distance_pred <- exp(predict(bee_ord_distance, newdata = preddat))
preddat$propCan_pred <- predict(bee_ord_propCan, newdata = preddat)

preddat <- preddat %>% 
  mutate(propCan_pred = case_when(
    propCan_pred < 0 ~ 0,
    TRUE ~ propCan_pred
  ))

bee_species <- data.frame(scores(bee_ord)$species) %>% 
  rownames_to_column("genus")

bee_species$distance_centroid <- exp(predict(
  bee_ord_distance, newdata = bee_species))
bee_species$propCan_centroid <- predict(
  bee_ord_propCan, newdata = bee_species)

bee_species <- bee_species %>% 
  left_join(., reduced_bees, by = "genus")


plot_bee_distance <- ggplot(mapping = aes(x = NMDS1, y = NMDS2)) +
  geom_contour(data = preddat,
               aes(z = distance_pred), color = "grey") +
  metR::geom_text_contour(data = preddat,
                          aes(z = distance_pred),
                          color = "grey", stroke = 0.15, skip = 0) +
  geom_text_repel(data = bee_species, max.overlaps = 10,
                  aes(label = genus, size = total_counts)) +
  scale_size_binned(transform = "log10", range = c(1.5, 5),
             breaks = c(3, 30, 300)) +
  theme_cowplot() +
  stat_ellipse(data = bee_sites,
               aes(x = NMDS1, y = NMDS2, color = period)) +
  geom_point(data = bee_sites_centroids,
             aes(x = NMDS1, y = NMDS2, fill = period),
             shape = 21, size = 5, show.legend = F) +
  # geom_point(data = bee_sites,
  #            aes(x = NMDS1, y = NMDS2, color = period),
  #            show.legend = F) +
  labs(size = "Abundance",
       color = "Season") +
  scale_color_manual(values = c("orange2",
                                "maroon3",
                                "seagreen3")) +
  scale_fill_manual(values = c("orange2",
                                "maroon3",
                                "seagreen3"))

plot_bee_propCan <- ggplot(mapping = aes(x = NMDS1, y = NMDS2)) +
  geom_contour(data = preddat,
               aes(z = propCan_pred), color = "grey") +
  metR::geom_text_contour(data = preddat,
                          aes(z = propCan_pred),
                          color = "grey", stroke = 0.15, skip = 1) +
  geom_text_repel(data = bee_species, max.overlaps = 20,
                  aes(label = genus, size = total_counts)) +
  scale_size_binned(transform = "log10", range = c(1.5, 5),
             breaks = c(3, 30, 300)) +
  theme_cowplot() +
  stat_ellipse(data = bee_sites,
               aes(x = NMDS1, y = NMDS2, color = period)) +
  geom_point(data = bee_sites_centroids,
             aes(x = NMDS1, y = NMDS2, fill = period),
             shape = 21, size = 5, show.legend = F) +
  # geom_point(data = bee_sites,
  #            aes(x = NMDS1, y = NMDS2, color = period),
  #            show.legend = F) +
  labs(size = "Abundance",
       color = "Season") +
  scale_color_manual(values = c("orange2",
                                "maroon3",
                                "seagreen3")) +
  scale_fill_manual(values = c("orange2",
                                "maroon3",
                                "seagreen3"))

cowplot::plot_grid(plot_bee_distance,
          plot_bee_propCan,
          nrow = 2)

```


###

```{r bee_diversity_by_land_type, echo = F}

#####################################################
# matrix assembly
habitat_matrix <- bees %>% 
  mutate(period_site_land = paste(
    period, site, land_type, sep = "_")) %>% 
  group_by(genus, period_site_land) %>% 
  tally(n = "count") %>% 
  pivot_wider(names_from = genus,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("period_site_land")

habitat_meta <- bees %>% 
  select(land_type, period, site) %>% 
  distinct() %>% 
  mutate(period_site_land = factor(
    paste(period, site, land_type, sep = "_")))

```

```{r permanova_stats, echo = F, warning=FALSE}
# land_type
land_type_mod <- adonis2(habitat_matrix ~ land_type * period,
        data = habitat_meta, method = "bray")

land_type_mod

permutest(betadisper(vegdist(habitat_matrix, method = "bray"),
                     habitat_meta$land_type))

# bloom period
period_mod <- adonis2(habitat_matrix ~ period,
        data = habitat_meta, method = "bray")

period_mod

permutest(betadisper(vegdist(habitat_matrix, method = "bray"),
                     habitat_meta$period))

# site
site_mod <- adonis2(habitat_matrix ~ site,
        data = habitat_meta, method = "bray")

site_mod

permutest(betadisper(vegdist(habitat_matrix, method = "bray"),
                     habitat_meta$site))

```


### Site classification by species composition (Random Forest)
```{r random_forest_class, echo = F}

rf_matrix <- habitat_matrix %>% 
  rownames_to_column("period_site_land") %>% 
  separate_wider_delim(period_site_land, delim = "_",
                       cols_remove = F,
                       names = c("period", "site", "land_type")) %>% 
  mutate(across(c(period : period_site_land), as.factor))

rfmod <- randomForest(y = rf_matrix$land_type,
                           x = select(rf_matrix,
                                      -period, -site,
                                      -period_site_land),
                           importance = T, nPerm = 999,
                           proximity = T)


train(y = rf_matrix$land_type,
      x = select(rf_matrix, -period, -site,
                 -period_site_land),
      method = "rf", na.action = na.omit)

update(rfmod, mtry = 2)

importance(rfmod)
imp_vals_top <- data.frame(importance(rfmod)) %>% 
  arrange(desc(MeanDecreaseGini)) %>% 
  filter(MeanDecreaseGini > quantile(MeanDecreaseGini, 0.8))

kable(imp_vals_top, format = "markdown", digits = 3)

rfmod

MDSplot(rfmod, rf_matrix$land_type)
```

**Random Forest classification of habitat type by species composition.** The model was tuned without pre-processing. Overall model accuracy was 81%. The model can delineate the habitat types with an overall *out of bag error* of 3.33% using bee community composition as the predictor matrix. 




```{r nmds_sites_calcs, eval = T, include = F}
# NMDS scores
nmds_ltp <- metaMDS(habitat_matrix, distance = "bray",
                    k = 2, try = 250, autotransform = F)
nmds_ltp$stress
# 
stressplot(nmds_ltp)

ltp_scores <- data.frame(scores(nmds_ltp)$sites) %>%
  rownames_to_column("period_site_land") %>% 
  mutate(period_site_land = factor(period_site_land)) %>% 
  left_join(habitat_meta, by = "period_site_land") 

spp_scores <- data.frame(scores(nmds_ltp)$species) %>% 
  rownames_to_column("Species")

centroids_ltp <- ltp_scores %>% 
  group_by(period_site_land) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

centroids_habitat <- ltp_scores %>% 
  group_by(land_type) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

centroids_period <- ltp_scores %>% 
  group_by(period) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))
```

### Bee composition by habitat type
```{r nmds_habitat_fig, echo = F}

imp_spp <- imp_vals_top %>% 
  rownames_to_column("Species") %>% 
  select(Species, MeanDecreaseGini)

spp_scores <- right_join(spp_scores, imp_spp, by = "Species")

spp_scores <- spp_scores %>% 
  filter(!Species == "land_type")

ltp_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(NMDS1, NMDS2, color = land_type),
             size = 1) +
  stat_ellipse(aes(color = land_type),
               show.legend = F) +
  geom_point(data = centroids_habitat,
             aes(x = NMDS1, y = NMDS2, fill = land_type),
             size = 4,
             shape = 21, show.legend = F) +
  theme_cowplot() +
  ggrepel::geom_text_repel(data = spp_scores,
                           aes(x = NMDS1, y = NMDS2,
                               label = Species,
                               size = MeanDecreaseGini),
                           max.overlaps = 20)+
  scale_size_continuous(
    name = "      Mean\ndecrease Gini",
    limits = c(.4, 3),
    breaks = c(0.75, 2, 3),
    range = c(1, 6)) +
  guides(color = guide_legend(override.aes = list(size = 4, shape = 16))) +
  scale_color_manual(values = c("#E69F00", "#56B4E9"),
                     name = "Habitat type")+
  scale_fill_manual(values = c("#E69F00", "#56B4E9"))



```
**Variation in community composition across habitat types.** Bee species are plotted on a two-dimensional non-metric multidimensional scaling ordination of the 30 combinations of site, period, and habitat type. Small points are the individual site/period/habitat combinations. Large points are the centroids of the two habitat types with natural habitat in orange, semi-natural habitat in blue. Ellipses are 95% confidence intervals around the habitat centroids. Bees species shown are the most representative (top 20th percentile of a random forest analysis) of the compositional differences among sites. Text size of the labels is proportional to variable importance score (mean decrease in Gini score).



### Bee composition by habitat type and bloom period
```{r nmds_period_habitat_fig, echo = F}

ltp_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(NMDS1, NMDS2, color = period),
             size = 1) +
  stat_ellipse(aes(color = period),
               show.legend = F) +
  geom_point(data = centroids_period,
             aes(x = NMDS1, y = NMDS2, fill = period),
             size = 4,
             shape = 21, show.legend = F) +
  theme_cowplot() +
  ggrepel::geom_text_repel(data = spp_scores,
                           aes(x = NMDS1, y = NMDS2,
                               label = Species,
                               size = MeanDecreaseGini),
                           max.overlaps = 20)+
  scale_size_continuous(
    name = "      Mean\ndecrease Gini",
    limits = c(.4, 3),
    breaks = c(0.75, 2, 3),
    range = c(1, 6)) +
  guides(color = guide_legend(override.aes = list(size = 4, shape = 16))) +
  scale_color_manual(values = c("#E69F00", "#56B4E9", "#009E73"),
                     name = "Bloom")+
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73")) +
  facet_wrap(~land_type) 

```

Note that there is considerable overlap in community composition at all comparisons (PERMANOVA **habitat;** *P* = 0.44, *F* = 0.99, **site;** *P* = 0.27, *F* = 1.12, **period**; *P* = 0.54, *F* = 0.94). From this analysis, we can conclude that community bee community composition does not vary across natural/semi-natural habitat types, sites, or bloom period.




## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


