---
title: "Canola Bee Analysis"
author: "Dr. Riley M. Anderson, Olivia Shaffer, & Salena Helmreich"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

This analysis explores Salena and Olivia's canola experiment.

### Summary of Results

* No difference in bee community composition across habitat types (natural or semi-natural), or habitat types and bloom period.

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(glmmTMB)
library(lme4)
library(sjPlot)
library(car)
library(knitr)
library(emmeans)
library(vegan)
library(randomForest)
library(caret)


# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
bees <- read.csv("data/Bee_Samples_2022.csv")

plants <- read.csv("data/PlantDiv2022.csv")

bee_hw <- read.csv("data/CleanPalouseBees.csv")

###### Functions


```


```{r Data_Wrangling, echo = F, comment = ""}

# clean up Salena's data:
bee_hw <- bee_hw %>% 
  mutate(across(
    c(period : land_type), factor),
    across(c(transect : genus), factor),
    date = ymd(date),
    week = week(date),
    site = factor(case_when(
      site == "Smoot " ~ "Smoot",
      TRUE ~ site
    )),
    period = factor(case_when(
      period == "1" ~ "Pre-Bloom",
      period == "2" ~ "Peak-Bloom",
      TRUE ~ "Post-Bloom"
    )))

# Clean up Olivia's data:
bees <- bees %>% 
  select(-X) %>% 
  mutate(across(c(Period : Genus), factor),
         Date = ymd(Date),
         week = week(Date),
         Family = factor(case_when(
           Family == "Halictus " ~ "Halictidae",
           Family == "Halictidae" & Genus == "Nomada" ~ "Apidae",
           TRUE ~ Family
         )),
         Genus = factor(case_when(
           Genus == "Ceratina " ~ "Ceratina",
           Genus == "Melissodes " ~ "Melissodes",
           TRUE ~ Genus
         ))) %>% 
  rename(date = Date,
         period = Period,
         site = Site,
         transect = Transect,
         morphogroup = Morphogroup,
         family = Family,
         genus = Genus)

# Extract Salena's land type variable:
land_type <- bee_hw %>% 
  select(site, land_type) %>% 
  distinct()

# Add land type to Olivia's data:
bees <- bees %>% 
  left_join(land_type, by = "site")

#######################################################

# Which genera are found in both natural and semi-natural sites?

# nat <- bees1 %>% 
#   filter(land_type == "natural") %>% 
#   select(genus) %>% 
#   distinct()
# 
# seminat <- bees1 %>% 
#   filter(land_type == "semi") %>% 
#   select(genus) %>% 
#   distinct()
# 
# # species only found in NATURAL sites:
# bees_unique_natural <- anti_join(nat, seminat)
# 
# # species only found in SEMINATURAL sites:
# bees_unique_semi <- anti_join(seminat, nat)
# 
# # overlapping species (species we can use in site comparison models)
# overlap <- semi_join(nat, seminat, by = "genus")
# 
# # removing species that were only found in either natural or seminatural sites:
# bees.reduced <- semi_join(bees1, overlap, by = "genus")
# 
# 
# # removing genera with sample size < 3 in either land type:
# good_samples <- bees.reduced %>% 
#   group_by(land_type, genus) %>% 
#   tally() %>% 
#   filter(n > 2) %>% 
#   ungroup() %>% 
#   select(genus) %>% 
#   distinct()
# 
# 
# bees.reduced <- semi_join(bees.reduced, good_samples, by = "genus")



```

## Bee community composition

```{r bee_diversity_by_land_type, echo = F}

#####################################################
# matrix assembly
habitat_matrix <- bees %>% 
  mutate(period_site_land = paste(
    period, site, land_type, sep = "_")) %>% 
  group_by(genus, period_site_land) %>% 
  tally(n = "count") %>% 
  pivot_wider(names_from = genus,
              values_from = count,
              values_fill = list(count = 0)) %>% 
  column_to_rownames("period_site_land")

habitat_meta <- bees %>% 
  select(land_type, period, site) %>% 
  distinct() %>% 
  mutate(period_site_land = factor(
    paste(period, site, land_type, sep = "_")))

```

```{r permanova_stats, echo = F, warning=FALSE}
# land_type
land_type_mod <- adonis2(habitat_matrix ~ land_type,
        data = habitat_meta, method = "bray")

land_type_mod

permutest(betadisper(vegdist(habitat_matrix, method = "bray"),
                     habitat_meta$land_type))

# bloom period
period_mod <- adonis2(habitat_matrix ~ period,
        data = habitat_meta, method = "bray")

period_mod

permutest(betadisper(vegdist(habitat_matrix, method = "bray"),
                     habitat_meta$period))

# site
site_mod <- adonis2(habitat_matrix ~ site,
        data = habitat_meta, method = "bray")

site_mod

permutest(betadisper(vegdist(habitat_matrix, method = "bray"),
                     habitat_meta$site))

```


### Site classification by species composition (Random Forest)
```{r random_forest_class, echo = F}

rf_matrix <- habitat_matrix %>% 
  rownames_to_column("period_site_land") %>% 
  separate_wider_delim(period_site_land, delim = "_",
                       cols_remove = F,
                       names = c("period", "site", "land_type")) %>% 
  mutate(across(c(period : period_site_land), as.factor))

rfmod <- randomForest(y = rf_matrix$land_type,
                           x = select(rf_matrix,
                                      -period, -site,
                                      -period_site_land),
                           importance = T, nPerm = 999,
                           proximity = T)


train(y = rf_matrix$land_type,
      x = select(rf_matrix, -period, -site,
                 -period_site_land),
      method = "rf", na.action = na.omit)

update(rfmod, mtry = 2)

importance(rfmod)
imp_vals_top <- data.frame(importance(rfmod)) %>% 
  arrange(desc(MeanDecreaseGini)) %>% 
  filter(MeanDecreaseGini > quantile(MeanDecreaseGini, 0.8))

kable(imp_vals_top, format = "markdown", digits = 3)

# MDSplot(rfmod, rf_matrix$land_type)
```

**Random Forest classification of habitat type by species composition.** The model was tuned without pre-processing. Overall model accuracy was 81%. The model can delineate the habitat types with an overall *out of bag error* of 3.33% using bee community composition as the predictor matrix. 




```{r nmds_sites_calcs, eval = T, include = F}
# NMDS scores
nmds_ltp <- metaMDS(habitat_matrix, distance = "bray",
                    k = 2, try = 250, autotransform = F)
nmds_ltp$stress
# 
stressplot(nmds_ltp)

ltp_scores <- data.frame(scores(nmds_ltp)$sites) %>%
  rownames_to_column("period_site_land") %>% 
  mutate(period_site_land = factor(period_site_land)) %>% 
  left_join(habitat_meta, by = "period_site_land") 

spp_scores <- data.frame(scores(nmds_ltp)$species) %>% 
  rownames_to_column("Species")

centroids_ltp <- ltp_scores %>% 
  group_by(period_site_land) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

centroids_habitat <- ltp_scores %>% 
  group_by(land_type) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))

centroids_period <- ltp_scores %>% 
  group_by(period) %>% 
  summarise(NMDS1 = mean(NMDS1),
            NMDS2 = mean(NMDS2))
```

### Bee composition by habitat type
```{r nmds_habitat_fig, echo = F}

imp_spp <- imp_vals_top %>% 
  rownames_to_column("Species") %>% 
  select(Species, MeanDecreaseGini)

spp_scores <- right_join(spp_scores, imp_spp, by = "Species")

spp_scores <- spp_scores %>% 
  filter(!Species == "land_type")

ltp_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(NMDS1, NMDS2, color = land_type),
             size = 1) +
  stat_ellipse(aes(color = land_type),
               show.legend = F) +
  geom_point(data = centroids_habitat,
             aes(x = NMDS1, y = NMDS2, fill = land_type),
             size = 4,
             shape = 21, show.legend = F) +
  theme_cowplot() +
  ggrepel::geom_text_repel(data = spp_scores,
                           aes(x = NMDS1, y = NMDS2,
                               label = Species,
                               size = MeanDecreaseGini),
                           max.overlaps = 20)+
  scale_size_continuous(
    name = "      Mean\ndecrease Gini",
    limits = c(.4, 3),
    breaks = c(0.75, 2, 3),
    range = c(1, 6)) +
  guides(color = guide_legend(override.aes = list(size = 4, shape = 16))) +
  scale_color_manual(values = c("#E69F00", "#56B4E9"),
                     name = "Habitat type")+
  scale_fill_manual(values = c("#E69F00", "#56B4E9"))



```
**Variation in community composition across habitat types.** Bee species are plotted on a two-dimensional non-metric multidimensional scaling ordination of the 30 combinations of site, period, and habitat type. Small points are the individual site/period/habitat combinations. Large points are the centroids of the two habitat types with natural habitat in orange, semi-natural habitat in blue. Ellipses are 95% confidence intervals around the habitat centroids. Bees species shown are the most representative (top 20th percentile of a random forest analysis) of the compositional differences among sites. Text size of the labels is proportional to variable importance score (mean decrease in Gini score).



### Bee composition by habitat type and bloom period
```{r nmds_period_habitat_fig, echo = F}

ltp_scores %>% 
  ggplot(aes(x = NMDS1, y = NMDS2)) +
  geom_point(aes(NMDS1, NMDS2, color = period),
             size = 1) +
  stat_ellipse(aes(color = period),
               show.legend = F) +
  geom_point(data = centroids_period,
             aes(x = NMDS1, y = NMDS2, fill = period),
             size = 4,
             shape = 21, show.legend = F) +
  theme_cowplot() +
  ggrepel::geom_text_repel(data = spp_scores,
                           aes(x = NMDS1, y = NMDS2,
                               label = Species,
                               size = MeanDecreaseGini),
                           max.overlaps = 20)+
  scale_size_continuous(
    name = "      Mean\ndecrease Gini",
    limits = c(.4, 3),
    breaks = c(0.75, 2, 3),
    range = c(1, 6)) +
  guides(color = guide_legend(override.aes = list(size = 4, shape = 16))) +
  scale_color_manual(values = c("#E69F00", "#56B4E9", "#009E73"),
                     name = "Bloom")+
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73")) +
  facet_wrap(~land_type) 

```

Note that there is considerable overlap in community composition at all comparisons (PERMANOVA **habitat;** *P* = 0.44, *F* = 0.99, **site;** *P* = 0.27, *F* = 1.12, **period**; *P* = 0.54, *F* = 0.94). From this analysis, we can conclude that community bee community composition does not vary across natural/semi-natural habitat types, sites, or bloom period.




## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


