---
title: "Canola Analysis"
author: "Dr. Riley M. Anderson & Salena Helmreich"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

This analysis explores Salena's canola experiment.

### Summary of Results

* Some bees have reduced head width in semi natural areas while others have increased head width in semi natural areas.

* Most bees have head widths that vary independently of land type.

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(glmmTMB)
library(lme4)
library(sjPlot)
library(car)
library(knitr)
library(emmeans)

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
bees1 <- read.csv("data/CleanPalouseBees.csv")
# Salena: I manually replaced the Halictus Family with Halictidae in the data
# before importing it. It is saved this way with the correction.
# Also corrected one record of Lasioglossum that was incorrectly assigned to 
# the Apidae family.

# import the buffer data
km1 <- read.csv("data/canola_1kmbuff.csv")

km2 <- read.csv("data/canola_2kmbuff.csv")



###### Functions

# Function to return a list with unscaled labels and scaled breaks
# For plotting scaled & centered data
unscale.plot.axis <- function(var, center = TRUE, labels = NULL){
  
    # Check if labels are specified. If not, create them
    if (is.null(labels)) {
      
        # Create 6 labels from the minimum to the maximum of var
        labels <- seq(min(var), max(var), 
            by = floor((max(var) - min(var)) / 5))
    }
  
    # Make a list to hold the output
    out <- list()
    
    # Add the labels to the list
    out$labels <- labels
    
    # Calculate the breaks on the scaled scale using the scale and center
    out$breaks <- (out$labels - 
        attributes(scale(var, center = center))$`scaled:center`) / 
        attributes(scale(var, center = center))$`scaled:scale`
    
    # Return the list with labels and breaks
    return(out)
}

#############

shift_legend <- function(p){

  # check if p is a valid object
  if(!"gtable" %in% class(p)){
    if("ggplot" %in% class(p)){
      gp <- ggplotGrob(p) # convert to grob
    } else {
      message("This is neither a ggplot object nor a grob generated from ggplotGrob. Returning original plot.")
      return(p)
    }
  } else {
    gp <- p
  }

  # check for unfilled facet panels
  facet.panels <- grep("^panel", gp[["layout"]][["name"]])
  empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]]))
  empty.facet.panels <- facet.panels[empty.facet.panels]
  if(length(empty.facet.panels) == 0){
    message("There are no unfilled facet panels to shift legend into. Returning original plot.")
    return(p)
  }

  # establish extent of unfilled facet panels (including any axis cells in between)
  empty.facet.panels <- gp[["layout"]][empty.facet.panels, ]
  empty.facet.panels <- list(min(empty.facet.panels[["t"]]), min(empty.facet.panels[["l"]]),
                             max(empty.facet.panels[["b"]]), max(empty.facet.panels[["r"]]))
  names(empty.facet.panels) <- c("t", "l", "b", "r")

  # extract legend & copy over to location of unfilled facet panels
  guide.grob <- which(gp[["layout"]][["name"]] == "guide-box")
  if(length(guide.grob) == 0){
    message("There is no legend present. Returning original plot.")
    return(p)
  }
  gp <- gtable_add_grob(x = gp,
                        grobs = gp[["grobs"]][[guide.grob]],
                        t = empty.facet.panels[["t"]],
                        l = empty.facet.panels[["l"]],
                        b = empty.facet.panels[["b"]],
                        r = empty.facet.panels[["r"]],
                        name = "new-guide-box")

  # squash the original guide box's row / column (whichever applicable)
  # & empty its cell
  guide.grob <- gp[["layout"]][guide.grob, ]
  if(guide.grob[["l"]] == guide.grob[["r"]]){
    gp <- gtable_squash_cols(gp, cols = guide.grob[["l"]])
  }
  if(guide.grob[["t"]] == guide.grob[["b"]]){
    gp <- gtable_squash_rows(gp, rows = guide.grob[["t"]])
  }
  gp <- gtable_remove_grobs(gp, "guide-box")

  return(gp)
}

```


```{r Data_Wrangling, echo = F, comment = ""}

# look at the structure of your data:
# str(bees1)

# convert your character variables into factors:
bees1 <- bees1 %>% 
  # this code just uses less verbiage than individually mutating them
  mutate(across(c(period:land_type), factor),
         across(c(transect: genus), factor),
         date = ymd(date)) 


# add any other data wrangling code to this chunk


##################################################

# Which genera are found in both natural and semi-natural sites?

nat <- bees1 %>% 
  filter(land_type == "natural") %>% 
  select(genus) %>% 
  distinct()

seminat <- bees1 %>% 
  filter(land_type == "semi") %>% 
  select(genus) %>% 
  distinct()

# species only found in NATURAL sites:
bees_unique_natural <- anti_join(nat, seminat)

# species only found in SEMINATURAL sites:
bees_unique_semi <- anti_join(seminat, nat)

# overlapping species (species we can use in site comparison models)
overlap <- semi_join(nat, seminat, by = "genus")

# removing species that were only found in either natural or seminatural sites:
bees.reduced <- semi_join(bees1, overlap, by = "genus")


# removing genera with sample size < 3 in either land type:
good_samples <- bees.reduced %>% 
  group_by(land_type, genus) %>% 
  tally() %>% 
  filter(n > 2) %>% 
  ungroup() %>% 
  select(genus) %>% 
  distinct()


bees.reduced <- semi_join(bees.reduced, good_samples, by = "genus")



#####
# buffer data
km2 <- km2 %>% 
  select(Name, buffer2km)


buffer <- left_join(km1, km2, by = "Name")

buffer <- buffer %>% 
  rename(site = Name) %>% 
  mutate(site = factor(site))

bees.reduced <- left_join(bees.reduced, buffer, by = "site")

```

## Exploratory data analysis

* Head width by genus, family, and land type
```{r exp_bee_health_by_land_type_family_genus, echo = F}
# graph name ^ (use underscores for consistency)

# Put each graph in a separate code block with a different name.
# The title of each graph block will be automatically saved as the image name.
# The set up chunk earlier defines this and automatically generates a 
# .png & .pdf of every graph (so long as each chunk as only one graph)

ggplot(bees.reduced, aes(x = genus,
                         y = head_width,
                         fill = land_type)) +
  geom_boxplot() +
  facet_wrap(~family, scales = "free_y", nrow = 3) +
  theme_classic() +
  labs(x = "Genus", y = "Head width (mm)", fill = "Habitat") +
  coord_flip() 



```

* Intertegular space by genus, family, and land type
```{r intertegularspace_landtype_Family, echo = F}
ggplot(bees.reduced,
       aes(x = genus, y = intertegular_space, fill = land_type)) +
  geom_boxplot() +
  facet_wrap(~family, scales = "free_y", nrow = 3) +
  theme_classic(base_size = 14) +
  labs(x = "Genus", y = "Intertegular space (mm)", fill = "Habitat") +
  coord_flip() 
  


```

* Correlation between head width and intertegular space
```{r corrplot_headwidth_intertegularspace, echo = F}


ggplot(bees1, aes(x = head_width, y = intertegular_space)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  geom_text(x = 2.5, y = 4.2,
            label = round(
              cor(bees1$head_width,
                  bees1$intertegular_space,
                  method = "pearson"),
              3)
            ) +
  geom_text(x = 2.2, y = 4.2,
            label = "r = ") +
  theme_classic()

coors <- cor.test(bees1$head_width,
                  bees1$intertegular_space,
                  method = "pearson")

```

* Head width by morphogroup and land type
```{r headwidth_landtype_morphogroup, echo = F}

ggplot(bees.reduced,
       aes(x = morphogroup, y = head_width, fill = land_type)) +
  geom_boxplot() +
  theme_classic() +
  labs(x = "Morphogroup", y = "Head width (mm)", fill = "Habitat")

```

* Histogram of head width
```{r histogram_headwidth, echo = F}

ggplot(bees.reduced,
       aes(x = head_width, fill = land_type)) +
  geom_histogram(bins = 50) +
  facet_wrap(~genus, scales = "free") +
  theme_classic() +
  labs(x = "Head width (mm)", y = "Frequency", fill = "Habitat")

```

* Density plot of head width
```{r density_plot_headwidth,, echo = F}

ggplot(bees.reduced,
       aes(x = head_width, fill = land_type)) +
  geom_density(alpha = 0.7) +
  facet_wrap(~genus, scales = "free") +
  theme_classic() +
  labs(x = "Head width (mm)", y = "Density", fill = "Habitat")

```

Only a few genera have useful replication: Andrena, Ceratina, Halictus, Lasioglossum, Nomada, Osmia, and Sphecodes. Interestingly, Halictus appears very bimodal, perhaps there are two distinct species that we could separate. 

## Models

### Head width ~ genus x habitat type
```{r head_width_model1, echo = F}

# remove the genera that wont be useful for the analysis:
model_bees <- bees.reduced %>% 
  filter(genus == "Andrena" | genus == "Ceratina" | genus == "Halictus"
         | genus == "Lasioglossum" | genus == "Nomada" | 
           genus == "Osmia" | genus == "Sphecodes")


mod1 <- lmer(head_width ~ land_type * genus + transect +  period + (1 | site),
             contrasts = list(transect = "contr.sum",
                              period = "contr.sum"),
                data = model_bees)


plot_model(mod1, type = "pred",
           terms = c("genus","land_type")) +
  theme_classic(base_size = 14) +
  labs(x = "", y = "Head width (mm)", color = "Habitat",
       title = "") +
  geom_text(aes(x = 1.3, y = 2.7, label = "*"),
            size = 6,, color = 'black') + 
  geom_text(aes(x = 3.3, y = 2.2, label = "**"),
            size = 6,, color = 'black') +
    geom_text(aes(x = 6.3, y = 2.45, label = "**"),
            size = 6,, color = 'black') +
  coord_flip()



```

**Figure XX. Bee head width by genus and habitat type.** Points are the predicted values from a linear mixed model. Whiskers are 95% confidence intervals. Asterisks indicate significant habitat type effects at: * *P* < 0.1, ** *P* < 0.05.



```{r model_results, echo = F}

Anova(mod1)
summary(mod1)

emm_mod1 <- emmeans(mod1, ~ land_type | genus)


mod.cons <- as.data.frame(contrast(emm_mod1, "pairwise"))


##########################################
# counts for results:

# number of genera (total)
num.genera <- bees1 %>% select(genus) %>% distinct() %>% tally()

# number of families (total)
num.family <- bees1 %>% select(family) %>% distinct() %>% tally()



```

* Stats methods

To increase parity in our analysis, We restricted our records to include only bee genera that occurred in both natural and semi-natural habitats and had at least 3 records in each habitat type. Consistent with previous studies (citations), bee head width and intertegular space were highly correlated (*r* = 0.95, *P* < 0.001, 2). Therefore, we haphazardly opted to use head width exclusively in our analyses. Data preprocessing was performed with the *tidyverse* ecosystem of packages (Wickham et al. 2019).

To test the prediction that semi-natural habitats would reduce bee health, we modeled bee head width (our proxy for bee health) as a function of habitat type using a linear mixed model with the lme4 package (Bates et al. 2015). Fixed effects included habitat type, genus, and their two-way interaction. We modeled the field collection site as a normally distributed random intercept. To account for variation in field collection transect, we opted to include this term as a fixed effect to avoid unreliable variance estimates as there were only 4 levels. We used an analysis of deviance to determine the contribution of each model term to the reduction in model deviance using the CAR package (Fox and Weisberg 2019). Simple slope post hoc tests were used to determine interaction effect sizes at the individual genus level using the EMmeans package (Lenth 2023). All analyses were conducted using R, version 4.2.3 (R Core Team 2023).

* Stats results

In total, we found `r nrow(bees1)` bees representing `r num.genera[1,1]` genera within `r num.family[1,1]` families. Bee abundance was biased towards semi-natural sites with `r bees1 %>% filter(land_type != "natural") %>% nrow()` bees found, while natural sites had `r bees1 %>% filter(land_type == "natural") %>% nrow()` bees. *you could add in something about the different species that were found in only the natural or semi natural sites*. 

Sample size and habitat overlap exclusions restricted our dataset to `r nrow(model_bees)` observations from seven genera; Andrena, Ceratina, Halictus, Lasioglossum, Nomada, Osmia, and Sphecodes. Overall, the habitat type and bee genera interaction significantly reduced model deviance ($\chi_{6}^{2}$ = `r round(Anova(mod1)[5,1], 2)`, *P* < 0.001). Post hoc tests revealed disparate effects of habitat type on head width for the various genera. Andrena and Halictus showed decreased head width in natural sites, relative to semi-natural habitats ($\beta_{Andrena}$ = `r round(mod.cons[1,3],2)` *P* = `r round(mod.cons[1,7],2)`; $\beta_{Halictus}$ = `r round(mod.cons[3,3],2)` *P* = `r round(mod.cons[3,7],2)`). Osmia showed the opposite, with increased head width in the natural habitat compared to the semi-natural habitat ($\beta_{Osmia}$ = `r round(mod.cons[6,3],2)` *P* = `r round(mod.cons[6,7],2)`). The head width of Ceratina, Lasioglossum, Nomada, and Sphecodes all varied independently of habitat type (*P* > 0.1) (Figure XX).


* Citations
```{r citations, echo = F}
citation("tidyverse")
citation("emmeans")
citation("lme4")
citation("car")
citation()

```

### Head width ~ habitat type * genus * season

* Don't include this in the chapter:
```{r head_width_model2, echo = F}

mod2 <- lmer(head_width ~ genus * land_type * period + transect +
               (1 | site), contrasts = list(transect = "contr.sum"),
             # Osmia and Sphecodes don't have enough data for this
             # comparison, with them, the model is rank deficient. 
             data = filter(model_bees, genus != "Sphecodes" &
                             genus != "Osmia"))

Anova(mod2)
summary(mod2)

plot_model(mod2,
           type = "pred",
           terms = c("period", "land_type", "genus")) +
  theme_classic() +
  labs(x = "Season",
       y = "Head width (mm)",
       color = "Habitat",
       title = "") +
  geom_line()

emm_mod2 <- emmeans(mod2, ~ land_type | genus * period)


contrast(emm_mod2, "pairwise")


```





```{r buffer_mods, echo = F, include = F}

# hist(buffer$buffer1km)
# hist(buffer$buffer2km)
# 
# bees.reduced %>% 
#   ggplot(aes(x = land_type, y = buffer2km)) +
#   geom_jitter() +
#   geom_boxplot()
# 
# bees.reduced %>% 
#   ggplot(aes(x = buffer2km, y = Head.width, color = Genus.x)) +
#   geom_point() + 
#   geom_smooth(method = 'lm') +
#   facet_wrap(~ Family.x)



# bees.reduced <- bees.reduced %>% 
#   mutate(s_canola = scale(buffer2km))
# 
# mod5 <- lmer(Head.width ~ s_canola * Morphogroup.x +
#                (s_canola | Genus.x), data = bees.reduced)
# 
# 
# summary(mod5)
# plot_model(mod5, type = "pred",
#            terms = c("s_canola","Morphogroup.x"))
# 
# plot_model(mod5, type = "re")

```

```{r Halictus_model, echo = F}

# halic.mod <- lm(Head.width ~ s_canola,
#                   data = filter(bees.reduced,
#                                 Genus.x == "Halictus"))
# 
# summary(halic.mod)

```

```{r halictus_plot, echo = F}

# bees.reduced %>% 
#   filter(Genus.x == "Halictus") %>% 
#   mutate(area1km = pi * 2000^1) %>% 
#   ggplot(aes(x = buffer2km/area1km, y = Head.width)) +
#   geom_point()

```


```{r Osmia_model, echo = F}
# 
# osmia.mod <- lm(Head.width ~ s_canola,
#                   data = filter(bees.reduced,
#                                 Genus.x == "Osmia"))
# 
# summary(osmia.mod)

```

```{r Nomada_model, echo = F}

# nomada.mod <- lm(Head.width ~ s_canola,
#                   data = filter(bees.reduced,
#                                 Genus.x == "Nomada"))
# 
# summary(nomada.mod)

```






```{r buffer2km_plot, echo = F}
# bees.reduced %>% 
#   filter(genus == c("Osmia", "Andrena", "Bombus", "Halictus")) %>% 
#   ggplot(aes(x = scale(buffer2km), y = head_width)) +
#   geom_point() +
#   geom_smooth(method = 'lm') +
#   facet_wrap(~ genus, scales = 'free') +
#   theme_classic()

```



## Are some species more common in natural areas?

* Don't include this in your chapter.

* WARNING
This is not a proper occupancy model as we aren't accounting for detection probabilities (uncertainty in detection efforts).

This is a quick glimpse at the odds of finding particular genera in semi natural areas relative to natural areas.

Proper occupancy modeling can be done (with a significant amount of work) if this is a question of interest.
```{r occupancy, echo = F}

# how many bees of each genera are there?
genera <- bees.reduced %>% 
  select(land_type, genus) %>% 
  group_by(land_type, genus) %>% 
  tally() %>% 
  rename(occupied = n) %>% 
  mutate(unoccupied = 
           if_else(
             grepl("natural", land_type),
             514 - occupied,
             498 - occupied
           ))

occ.mod <- glmer(cbind(occupied, unoccupied) ~
                   land_type + (land_type | genus),
                 family = binomial(),
                 data = genera)

summary(occ.mod)



```

```{r occupancy_fig, echo = F}

plot_model(occ.mod,
           type = "re") +
  theme_classic()


```

Again, the variance component for the random slope is small but there may be some trends for a few genera. For example, *Halictus* and *Ceratina* are more likely to be found in semi natural areas, while *Andrena* and *Osmia* are more likely to be found in natural areas. Keep in mind the simplified model structure here and its limitations. Also the model is run on a reduced data set including only genera that were found in both land types. With the full data set, there are more random effect levels than observations making the random effect parameters unidentifiable.


## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


