---
title: "2022FloralDiv"
author: "Olivia Shaffer"
date: "1/1/2023"
output: html_document
---

Based on:
https://rpubs.com/mbh038/719881

```{r}
install.packages("ggplot2")
install.packages("vegan")
install.packages("tidyverse")
install.packages("dplyr")
install.packages("lme4")
install.packages("lmerTest")
install.packages("multcompView")
```

```{r}
library("ggplot2")
library("vegan")
library("tidyverse")
library("dplyr")
library(lme4)
library(lmerTest)
library(multcompView)
```

load data
```{r}
flora <- read.csv("PlantDiv2022.csv")
write.csv(flora, "flora.csv")
```

Create new columns
```{r}
flora$Species <- paste(flora$Genus, flora$Species, sep="_")
flora$SiteSeason <- paste(flora$Site, flora$Period, sep="_")
```

Create interaction matrix
```{r}
#By site
BySite <- table(flora$Site, flora$Species)
as.matrix(BySite)

SiteSp <- table(flora$Species, flora$Site)
as.matrix(SiteSp)

#By Season
BySeason <- table(flora$Period, flora$Species)
as.matrix(BySeason)

SeasonSp <- table(flora$Species, flora$Period)
as.matrix(SeasonSp)


#SiteSeason
BySiteSeason<- table(flora$SiteSeason, flora$Species)
as.matrix(BySiteSeason)

SiteSeasSp <- table(flora$Species, flora$SiteSeason)
as.matrix(SiteSeasSp)

```

Create presence/absence tables for sp. richness
```{r}
#By Site
Sitepa<- decostand(SiteSp, method = "pa")

#By Season
Seasonpa <- decostand(SeasonSp, method = "pa")

#SiteSeason
SiteSeasonpa <- decostand(SiteSeasSp, method="pa")

#All
Allpa <-decostand(SiteSeasSp, method="pa")
```

#Total Sp Diversity
```{r}
#Species
SpeciesRichnessAll <- 

#Genera

#Family
```


#Diversity by Site

Plant Abundance
- total plants found in each sites
```{r}
#By Site
SiteA <-select(flora, Amount, Site)
#combine values 
SiteAbun<- SiteA %>%
  group_by(Site) %>%
  summarise(across(c(Amount), sum))

write.csv(SiteAbun, "AbundancebySitex.csv")

SiteAbundancedf <- data.frame(SiteAbun)
```

Species Richness
- num. of sp found in each sites
```{r}
SiteRichness <- colSums(Sitepa)

write.csv(SiteRichness, "SpRichnessbySite.csv")

SiteRichnessdf <- data.frame(SiteRichness)
```


Sp Diversity
```{r}
DivbySite <- diversity(BySite,"shannon")
write.csv(DivbySite, "DivBySite.csv")

SiteDivdf <- data.frame(DivbySite)
```


#Diversity by Season

Sp Abundance
```{r}

SeasonA <-select(flora, Amount, Period)

#combine values 
SeasonAbun <- SeasonA %>%
  group_by(Period) %>%
  summarise(across(c(Amount), sum))

write.csv(SeasonAbun, "AbundancebySeasonx.csv")

SeasonAbundancedf <- data.frame(SeasonAbun)
```

Species Richness
- num. of sp found in each season
```{r}
SeasRichness <- colSums(Seasonpa)

write.csv(SeasRichness, "SpRichnessbySeason.csv")

SeasRichnessdf <- data.frame(SeasRichness)
```

Sp Diversity
```{r}
DivbySeas <- diversity(BySeason,"shannon")
write.csv(DivbySeas, "DivBySeason.csv")
```

#Diversity by SiteSeason

Abundance
```{r}
#All
SiteSeasonA <-select(flora, Amount, SiteSeason)
#combine values 
SiteSeasonAbun <- SiteSeasonA %>%
  group_by(SiteSeason) %>%
  summarise(across(c(Amount), sum))

write.csv(SiteSeasonAbun, "AbundanceAll.csv")

SiteSeasonAbundancedf <- data.frame(SiteSeasonAbun)
```

Species Richness
- num. of sp found in each sites
```{r}
SiteSeasRichness <- colSums(SiteSeasonpa)

write.csv(SiteSeasRichness, "AllSpRichness.csv")

SiteSeasRichnessdf <- data.frame(SiteSeasRichness)
```


Diversity
```{r}
DivbyAll <- diversity(BySiteSeason,"shannon")
write.csv(DivbyAll, "AllDiv.csv")
```


Graphs!

Dissimilarity index 
-Determines whether species Diversity varies

Season
```{r}
Seas.dist<-vegdist(BySeason,binary=FALSE)

#standardize data
Seas.dist <- vegdist(decostand(BySeason, "norm"), "euclidean")

#Calc and plot
Seas.hc<-hclust(Seas.dist)
plot(Seas.hc)
```

Site
```{r}
Site.dist<-vegdist(BySite,binary=FALSE)

#standardize data
Site.dist <- vegdist(decostand(BySite, "norm"), "euclidean")

#Calc and plot
Site.hc<-hclust(Site.dist)
plot(Site.hc)
```



-graph showing Abundance & Richness
-graph showing diversity

-Created combined datasets in excel

Load in combined datasets
```{r}
AllSite <- read.csv("BySiteFloral.csv")
AllSite$Richness <- as.numeric(AllSite$Richness)

AllSite <- AllSite %>%
  select(Site, Richness, Shannon_Diversity) %>%
  gather(key = "variable", value = "value", -Site)

head(AllSite)

AllSeason <- read.csv("BySeasonFloral.csv")
AllSeason$Richness <- as.numeric(AllSeason$Richness)

AllSeason <- AllSeason %>%
  select(Season, Richness, Shannon_Diversity) %>%
  gather(key = "variable", value = "value", -Season)
head(AllSeason)

#to get seasons in the right order
AllSeason$Season <- as.character(AllSeason$Season)
AllSeason$Season <- factor(AllSeason$Season, levels=unique(AllSeason$Season))

AbundanceSeasonG <- read.csv("AbundancebySeasonx.csv")
AbundanceSeasonG$Period <- as.character(AbundanceSeasonG$Period)
AbundanceSeasonG$Period <- factor(AbundanceSeasonG$Period, levels=unique(AbundanceSeasonG$Period))

SiteAbun <- read.csv("AbundancebySitex.csv")
```


By Site:
```{r}
#Richness & Diversity
ggplot(AllSite, aes(fill= variable, x = Site, y=Amount, group=1)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  scale_fill_manual(values = c( '#FFC300', '#FF5733')) +
  ggtitle("Plant Diversity Values by Field Site") +
  theme(text = element_text(size = 20), axis.text.x = element_text(size = 15, angle = 90))

#Abundance
ggplot(SiteAbun, aes(x = Site, y=Amount, fill = '#900C3F', group=1)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  theme(legend.position = "none") +
  ggtitle("Plant Abundance by Field Site")+
  theme(text = element_text(size = 20), axis.text.x = element_text(size = 15, angle = 90))

```


By Season: 
```{r}
#Richness & Diversity
ggplot(AllSeason, aes(fill= variable, x = Season, y=value, group=1)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  scale_fill_manual(values = c( '#FFC300', '#FF5733')) +
  ggtitle("Plant Diversity Values by Sampling Period")

#Abundance
ggplot(AbundanceSeasonG, aes(x = Period, y=Amount, fill = '#900C3F', group=1)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  theme(legend.position = "none") +
  ggtitle("Plant Abundance by Sampling Period")
```

Testing for Sig Diff
```{r}
#Load in floral data by site & season
AllFloral <- read.csv("AllFloral.csv")


##Sites
#Abundance
PlantSiteAbun<- lm(Abundance~Site, data = AllFloral)
#P = 0.7736
report(PlantSiteAbun)

#Richness
PlantSiteRich <-lm(Richness~Site, data = AllFloral)
summary(lm(Richness~Site, data = AllFloral))
#P=0.1834
report(PlantSiteRich)

#Diversity
PlantSiteDiv <- lm(Shannon_Diversity~Site, data = AllFloral)
summary(lm(Shannon_Diversity~Site, data = AllFloral))
#P=0.0747
report(PlantSiteDiv)

##Seasons
#Abundance
PAbundanceSeas <- lm(Abundance~Season, data = AllFloral)
#P=0.0204
summary(aov(PAbundanceSeas))
PMixedAbunSite <- lmer(Abundance~Season + (1|Site), data = AllFloral)
summary(PMixedAbunSite)
report(PAbundanceSeas)

#Richness
PRichSeas <- lm(Richness~Season, data = AllFloral)
#P=0.0256
summary(aov(PRichSeas))
PMixedRichSite <- lmer(Richness~Season + (1|Site), data = AllFloral)
summary(PMixedRichSite)
report(PRichSeas)

#Diversity
PDivSeas <- lm(Shannon_Diversity~Season, data = AllFloral)
#P=0.0387
summary(aov(PDivSeas))
PMixedDivSeas <- lmer(Shannon_Diversity~Season + (1|Site), data = AllFloral)
summary(PMixedDivSeas)
report(PDivSeas)


#Fit of mixed v fixed effect models
anova(PMixedAbunSite, PAbundanceSite)
anova(PMixedRichSite, PRichSite)
anova(PMixedDivSite, PDivSite)
```

Tukey HSD

-By Season
```{r}
AllFloral <- read.csv("AllFloral.csv")

#Tukey Test
anova1 <- aov(Abundance~Season, data = AllFloral)
tukey1 <- TukeyHSD(anova1, conf.level=.95)
tukey1.cld <-multcompLetters4(anova1,tukey1)
summary(anova1)
print(tukey1)
print(tukey1.cld)

anova2 <- aov(Richness~Season, data = AllFloral)
tukey2 <- TukeyHSD(anova2, conf.level=.95)
tukey2.cld <-multcompLetters4(anova2,tukey2)
summary(anova2)
print(tukey2)
print(tukey2.cld)

anova3 <- aov(Shannon_Diversity~Season, data = AllFloral)
tukey3 <- TukeyHSD(anova3, conf.level=.95)
tukey3.cld <-multcompLetters4(anova3,tukey3)
summary(anova3)
print(tukey3)
print(tukey3.cld)

#Restructured dataset for error bars
FloralTukey <- read.csv("FloralTukeyTest.csv")

#Add tukey test results to dataset
Abundance.cld <- as.data.frame.list(tukey1.cld$'Season')
FloralTukey$Abundance_Tukey <- Abundance.cld$Letters 

Richness.cld <- as.data.frame.list(tukey2.cld$'Season')
FloralTukey$Richness_Tukey <- Richness.cld$Letters 

Diversity.cld <- as.data.frame.list(tukey3.cld$'Season')
FloralTukey$Diversity_Tukey <- Diversity.cld$Letters 

#Save restructured dataset
write.csv(FloralTukey, "FloralTukeyTest.csv")

#Plot
str(FloralTukey)
FloralTukey$Season <- as.factor(FloralTukey$Season)

#Abundance
ggplot(FloralTukey, aes(Season, Abundance_mean, fill =Abundance_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Set2")+
  labs(x = "Season", y = "Average Floral Abundance")+ 
  theme_bw()+
  scale_x_discrete(breaks=c("1PreBloom","2PeakBloom","3DeclineBloom"),
        labels=c("PreBloom", "PeakBloom", "DeclineBloom"))+
  geom_errorbar(aes(ymin=Abundance_mean-Abundance_SE, ymax=Abundance_mean+Abundance_SE), width = 0.2, color = "gray30")+ 
  geom_text(aes(label =Abundance_Tukey), nudge_y = 30, size = 5)+
  ggtitle("Average Floral Abundance by Season")

#Richness
ggplot(FloralTukey, aes(Season, Richness_mean, fill =Richness_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Paired")+
  labs(x = "Season", y = "Average Floral Richness")+ 
  theme_bw()+
  scale_x_discrete(breaks=c("1PreBloom","2PeakBloom","3DeclineBloom"),
        labels=c("PreBloom", "PeakBloom", "DeclineBloom"))+
  geom_errorbar(aes(ymin=Richness_mean-Richness_SE, ymax=Richness_mean+Richness_SE), width = 0.2, color = "gray30")+ 
  geom_text(aes(label =Richness_Tukey), nudge_y = 1.7, size = 5)+
  ggtitle("Average Floral Species Richness by Season")
  
#Diversity
ggplot(FloralTukey, aes(Season, Shannon_Diversity_mean, fill =Diversity_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Accent")+
  labs(x = "Season", y = "Average Floral Shannon Diversity")+ 
  theme_bw()+
  scale_x_discrete(breaks=c("1PreBloom","2PeakBloom","3DeclineBloom"),
        labels=c("PreBloom", "PeakBloom", "DeclineBloom"))+
  geom_errorbar(aes(ymin=Shannon_Diversity_mean-Shannon_Diversity_SE, ymax=Shannon_Diversity_mean+Shannon_Diversity_SE), width = 0.2, color = "gray30")+
  geom_text(aes(label =Diversity_Tukey), nudge_y = 0.4, size = 5)+
  ggtitle("Average Floral Shannon Diversity by Season")
```

-By Site
```{r}
#Tukey Test
anova4 <- aov(Abundance~Site, data = AllFloral)
tukey4 <- TukeyHSD(anova4, conf.level=.95)
tukey4.cld <-multcompLetters4(anova4,tukey4)
print(tukey4)
print(tukey4.cld)

anova5 <- aov(Richness~Site, data = AllFloral)
tukey5 <- TukeyHSD(anova5, conf.level=.95)
tukey5.cld <-multcompLetters4(anova5,tukey5)
print(tukey5)
print(tukey5.cld)

anova6 <- aov(Shannon_Diversity~Site, data = AllFloral)
tukey6 <- TukeyHSD(anova6, conf.level=.95)
tukey6.cld <-multcompLetters4(anova6,tukey6)
print(tukey6)
print(tukey6.cld)

#Restructured dataset for error bars
FloralSiteTukey <- read.csv("FloralSiteTukey.csv")

#Add tukey test results to dataset
Abundance.cld4 <- as.data.frame.list(tukey4.cld$'Site')
FloralSiteTukey$Abundance_Tukey <- Abundance.cld4$Letters 

Richness.cld5 <- as.data.frame.list(tukey5.cld$'Site')
FloralSiteTukey$Richness_Tukey <- Richness.cld5$Letters 

Diversity.cld6 <- as.data.frame.list(tukey6.cld$'Site')
FloralSiteTukey$Diversity_Tukey <- Diversity.cld6$Letters 

#Save restructured dataset
write.csv(FloralSiteTukey, "FloralSiteTukey.csv")


#Plot
str(FloralSiteTukey)
FloralSiteTukey$Site <- as.factor(FloralSiteTukey$Site)

#Abundance
ggplot(FloralSiteTukey, aes(Site, Abundance_mean, fill =Abundance_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Set2")+
  labs(x = "Site", y = "Average Floral Abundance")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Abundance_mean-Abundance_SE, ymax=Abundance_mean+Abundance_SE), width = 0.5, color = "gray30")+ 
  geom_text(aes(label =Abundance_Tukey), nudge_y = 140, size = 5)+
  ggtitle("Average Floral Abundance by Site")

#Richness
ggplot(FloralSiteTukey, aes(Site, Richness_mean, fill =Richness_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Paired")+
  labs(x = "Site", y = "Average Floral Richness")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Richness_mean-Richness_SE, ymax=Richness_mean+Richness_SE), width = 0.2, color = "gray30")+ 
  geom_text(aes(label =Richness_Tukey), nudge_y = 1.6, size = 5)+
  ggtitle("Average Floral Species Richness by Site")
  
#Diversity
ggplot(FloralSiteTukey, aes(Site, Shannon_mean, fill =Diversity_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Accent")+
  labs(x = "Site", y = "Average Floral Shannon Diversity")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Shannon_mean-Shannon_Diversity_SE, ymax=Shannon_mean+Shannon_Diversity_SE), width = 0.2, color = "gray30")+
  geom_text(aes(label =Diversity_Tukey), nudge_y = 0.6, size = 5)+
  ggtitle("Average Floral Shannon Diversity by Site")
```

Relationship with Canola

```{r}
Canola <- read.csv("CanolaPercentages.csv")
SiteFloral <- read.csv("BySiteFloral.csv")
```


Distance
```{r}
cor.test(Canola$DistancetoCanola, SiteFloral$Abundance)

ggplot() + 
  geom_point(mapping = aes(y = SiteFloral$Abundance, x= Canola$DistancetoCanola, color = Canola$Site)) + 
  geom_smooth(mapping = aes(y = SiteFloral$Abundance, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "Floral Abundance", x = "Distance to Canola", colour = "Site") +
  theme_bw()+
  ggtitle("Floral Abundance and Distance to Canola")

cor.test(Canola$DistancetoCanola, SiteFloral$Richness)

ggplot() + 
  geom_point(mapping = aes(y = SiteFloral$Richness, x= Canola$DistancetoCanola, color = Canola$Site)) + 
  geom_smooth(mapping = aes(y = SiteFloral$Richness, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "Floral Richness", x = "Distance to Canola", colour = "Site") +
  ggtitle("Floral Richness and Distance to Canola")

cor.test(Canola$DistancetoCanola, SiteFloral$Shannon_Diversity)

ggplot() + 
  geom_point(mapping = aes(y = SiteFloral$Shannon_Diversity, x= Canola$DistancetoCanola, color = Canola$Site)) + 
  geom_smooth(mapping = aes(y = SiteFloral$Shannon_Diversity, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "Floral Shannon Diversity", x = "Distance to Canola", colour = "Site") +
  ggtitle("Floral Diversity and Distance to Canola")
```

Percentage
```{r}
cor.test(Canola$Canola2km, SiteFloral$Abundance)

ggplot() + 
  geom_point(mapping = aes(y = SiteFloral$Abundance, x= Canola$X2kmCan., color = Canola$Site)) + 
  geom_smooth(mapping = aes(y = SiteFloral$Abundance, x = Canola$X2kmCan.), method = lm) +
  labs(y = "Floral Abundance", x = "Canola Percentage in 2km", colour = "Site") +
  ggtitle("Floral Abundance and Canola Percentage")

cor.test(Canola$Canola2km, SiteFloral$Richness)

ggplot() + 
  geom_point(mapping = aes(y = SiteFloral$Richness, x= Canola$X2kmCan., color = Canola$Site)) + 
  geom_smooth(mapping = aes(y = SiteFloral$Richness, x = Canola$X2kmCan.), method = lm) +
  labs(y = "Floral Richness", x = "Canola Percentage in 2km", colour = "Site") +
  ggtitle("Floral Richness and Canola Percentage")

cor.test(Canola$Canola2km, SiteFloral$Shannon_Diversity)

ggplot() + 
  geom_point(mapping = aes(y = SiteFloral$Shannon_Diversity, x= Canola$X2kmCan., color = Canola$Site)) + 
  geom_smooth(mapping = aes(y = SiteFloral$Shannon_Diversity, x = Canola$X2kmCan.), method = lm) +
  labs(y = "Floral Shannon Diversity", x = "Canola Percentage in 2km", colour = "Site") +
  ggtitle("Floral Diversity and Canola Percentage")
```


Species Histograms
```{r}
FlowerSp <- read.csv("flora.csv")

FlowerSp$Family <- as.factor(FlowerSp$Family)
str(FlowerSp)
```

Taxa composition
```{r}
ggplot(FlowerSp, aes(x= reorder(Family, -Amount), y = Amount, fill = Amount))+
  geom_bar(aes(fill = 'Family'), stat = "identity", position = position_stack())+ theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

ggplot(FlowerSp, aes(x= reorder(Genus, -Amount), y = Amount, fill = Amount))+
  geom_bar(aes(fill = 'Genus'), stat = "identity", position = position_stack())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

ggplot(FlowerSp, aes(x= reorder(Family, -Amount), y = Amount, fill = Amount))+
  geom_bar(aes(fill = 'Family'), stat = "identity", position = position_stack())+ theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

ggplot(FlowerSp, aes(x= reorder(Species, -Amount), y = Amount, fill = Amount))+
  geom_bar(aes(fill = 'Species'), stat = "identity", position = position_stack())+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

By Site
```{r}
ggplot(FlowerSp, aes(x= reorder(Site, -Amount), y = Amount, fill = Family))+
  geom_bar(stat = "identity", position = "stack")+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(x = "Site")+
  ggtitle("Floral Family by Site")
```

