---
title: "BeeDiv2022"
author: "Olivia Shaffer"
date: "1/6/2023"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Based on: <https://rpubs.com/mbh038/719881>



```{r}
library("ggplot2")
library("vegan")
library("tidyverse")
library("dplyr")
library(lme4)
library(lmerTest)
library(report)
library(emmeans)
library(multcompView)
```

Load in data

```{r}
bees <- read.csv("data/olivia_data/Bee_Samples_2022.csv")
```

Make SiteSeason column

```{r}
bees$SiteSeason <- paste(bees$Site, bees$Period, sep="_")
```

Create Interaction Matrices

```{r}
#By site
BySite <- table(bees$Site, bees$Genus)
as.matrix(BySite)

SiteGen <- table(bees$Genus, bees$Site)
as.matrix(SiteGen)

#By Season
BySeason <- table(bees$Period, bees$Genus)
as.matrix(BySeason)

SeasonGen <- table(bees$Genus, bees$Period)
as.matrix(SeasonGen)


#SiteSeason
BySiteSeason<- table(bees$SiteSeason, bees$Genus)
as.matrix(BySiteSeason)

SiteSeasGen <- table(bees$Genus, bees$SiteSeason)
as.matrix(SiteSeasGen)
```

Create presence/absence tables for sp. richness

```{r}
#By Site
Sitepa<- decostand(SiteGen, method = "pa")

#By Season
Seasonpa <- decostand(SeasonGen, method = "pa")

#All
SiteSeasonpa <- decostand(SiteSeasGen, method="pa")
```

#Diversity by Site

Bee Abundance - total bees found in each sites

```{r}
SiteAbun <- colSums(SiteGen)

write.csv(SiteAbun, "AbunbySiteB.csv")

SiteAbun <- as.data.frame(SiteAbun)
```

Species Richness - num. of sp found in each sites

```{r}
SiteRichness <- colSums(Sitepa)

write.csv(SiteRichness, "SpRichnessbySiteB.csv")

SiteRichnessdf <- data.frame(SiteRichness)
```

Sp Diversity

```{r}
DivbySite <- diversity(BySite,"shannon")
write.csv(DivbySite, "DivBySiteB.csv")

SiteDivdf <- data.frame(DivbySite)
```

#Diversity by Season

Sp Abundance

```{r}
SeasAbun <- colSums(SeasonGen)

write.csv(SeasAbun, "AbunbySeasB.csv")

SeasAbun <- as.data.frame(SeasAbun)
```

Species Richness - num. of sp found in each sites

```{r}
SeasRichness <- colSums(Seasonpa)

write.csv(SeasRichness, "SpRichnessbySeasonB.csv")

SeasRichnessdf <- data.frame(SeasRichness)
```

Sp Diversity

```{r}
DivbySeas <- diversity(BySeason,"shannon")
write.csv(DivbySeas, "DivBySeasonB.csv")
```

#Diversity by SiteSeason

Abundance

```{r}
SiteSeasAbun <- colSums(SiteSeasGen)

write.csv(SiteSeasAbun, "AbunbySiteSeasB.csv")

SiteSeasAbun <- as.data.frame(SiteSeasAbun)
```

Species Richness - num. of sp found in each sites

```{r}
SiteSeasRichness <- colSums(SiteSeasonpa)

write.csv(SiteSeasRichness, "AllSpRichness.csv")

SiteSeasRichnessdf <- data.frame(SiteSeasRichness)
```

Diversity

```{r}
DivbyAll <- diversity(BySiteSeason,"shannon")
write.csv(DivbyAll, "AllDiv.csv")
```

Graphs!

Dissimilarity index -Determines whether species Diversity varies

Season

```{r}
Seas.dist<-vegdist(BySeason,binary=FALSE)

#standardize data
Seas.dist <- vegdist(decostand(BySeason, "norm"), "euclidean")

#Calc and plot
Seas.hc<-hclust(Seas.dist)
plot(Seas.hc)
```

Site

```{r}
Site.dist<-vegdist(BySite,binary=FALSE)

#standardize data
Site.dist <- vegdist(decostand(BySite, "norm"), "euclidean")

#Calc and plot
Site.hc<-hclust(Site.dist)
plot(Site.hc)
```

#####Still needs edited from FloralDiv2022####

-graph showing Abundance & Richness -graph showing diversity

-Created combined datasets in excel

Load in combined datasets

```{r}
AllSite <- read.csv("data/olivia_data/BeesbySite.csv")
AllSite$Richness <- as.numeric(AllSite$Richness)

AllSite <- AllSite %>%
  select(Site, Richness, Shannon_Diversity) %>%
  gather(key = "variable", value = "value", -Site)

head(AllSite)

AllSeason <- read.csv("data/olivia_data/BeesbySeason.csv")
AllSeason$Richness <- as.numeric(AllSeason$Richness)

AllSeason <- AllSeason %>%
  select(Season, Richness, Shannon_Diversity) %>%
  gather(key = "variable", value = "value", -Season)
head(AllSeason)

##to get seasons in the right order
#AllSeason$Season <- as.character(AllSeason$Season)
#AllSeason$Season <- factor(AllSeason$Season, levels=unique(AllSeason$Season))

#AbundanceSeasonG <- read.csv("AbunbySeasB.csv")
#AbundanceSeasonG$Season <- as.character(AbundanceSeasonG$Season)
#AbundanceSeasonG$Season <- factor(AbundanceSeasonG$Season, levels=unique(AbundanceSeasonG$Season))
```

By Site:

```{r}
#Richness & Diversity
ggplot(AllSite, aes(fill= variable, x = Site, y=value, group=1)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  scale_fill_manual(values = c( '#FFC300', '#FF5733')) +
  ggtitle("Bee Diversity by Field Site")+
  theme(text = element_text(size = 20), axis.text.x = element_text(size = 15, angle = 90))


#Abundance
AbunbySite <- read.csv("data/olivia_data/AbunbySiteB.csv")

ggplot(AbunbySite, aes(x = Site, y=Abundance, fill = '#900C3F', group=1)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  theme(legend.position = "none") +
  ggtitle("Bee Abundance by Field Site")+
  theme(text = element_text(size = 20), axis.text.x = element_text(size = 15, angle = 90))

```

By Season:

```{r}
#Changing order of seasons. Above code not working for some reason
AllSeason$Season <- factor(AllSeason$Season, levels=c("Pre-Bloom", "Peak-Bloom", "Post-Bloom"))
AbundanceSeasonG$Season <- factor(AbundanceSeasonG$Season, levels=c("Pre-Bloom", "Peak-Bloom", "Post-Bloom"))

#Richness & Diversity
ggplot(AllSeason, aes(fill= variable, x = Season, y=value, group=1)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  scale_fill_manual(values = c( '#FFC300', '#FF5733')) +
  ggtitle("Bee Diversity by Sampling Period")

#Abundance
ggplot(AbundanceSeasonG, aes(x = Season, y=Abundance, fill = '#900C3F', group=1)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  theme(legend.position = "none") +
  ggtitle("Bee Abundance by Sampling Period")
```

Testing for Sig Diff

```{r}
#Load in bee data by site & season
AllBees <- read.csv("BeesbySiteSeason.csv")
#Corresponding floral data
AllFloral <- read.csv("AllFloral.csv")


##Sites
#Abundance
BeeAbunSite <- lm(Abundance~Site, data = AllBees)
summary(aov(BeeAbunSite))
summary(lm(Abundance~Site, data = AllBees))
#P = 0.6332
report(BeeAbunSite)

#Richness
BeeRichSite <- lm(Richness~Site, data = AllBees)
summary(aov(BeeRichSite))
summary(lm(Richness~Site, data = AllBees))
#P=0.009583
report(BeeRichSite)

#Diversity
BeeDivSite <- lm(Shannon_Diversity~Site, data = AllBees)
summary(aov(BeeDivSite))
summary(lm(Shannon_Diversity~Site, data = AllBees))
#P=0.00657
report(BeeDivSite)

##Seasons
#Abundance
AbundanceSeason <- lm(Abundance~Season, data = AllBees)
#P=0.1821
summary(aov(AbundanceSeason))
AbundanceSeasonMixed <- lmer(Abundance~Season +(1|Site),data=AllBees)
summary(AbundanceSeasonMixed)

report(AbundanceSeasonMixed)
#no sig diff between the two models. Report the 1st one
summary(AbundanceSeason)
report(AbundanceSeason)


AbundanceSite <- lmer(Abundance~ (1|Site),data=AllBees)


#Richness
RichSeason <- lm(Richness~Season, data = AllBees)
#P=0.7995
summary(aov(RichSeason))
RichSeasonMixed <- lmer(Richness~Season +(1|Site),data=AllBees)
summary(RichSeasonMixed)
report(RichSeason)

RichSite <- lmer(Richness~ (1|Site),data=AllBees)

#Diversity
DivSeason<- lm(Shannon_Diversity~Season, data = AllBees)
#P=0.8309
summary(aov(DivSeason))
DivSeasonMixed <- lmer(Shannon_Diversity~Season +(1|Site),data=AllBees)
summary(DivSeasonMixed)
report(DivSeason)
report(DivSeasonMixed)

DivSite <- lmer(Shannon_Diversity~ (1|Site),data=AllBees)

#Richness & Diversity vary by site


#Are mixed effect models better?
anova(AbundanceSeasonMixed, AbundanceSeason)
#simple model
anova(RichSeasonMixed, RichSeason)
#mixed better
anova(DivSeasonMixed, DivSeason)
#mixed better

#Should season be included or by site?
anova(AbundanceSite, AbundanceSeason)
anova(RichSite, RichSeason)
anova(DivSite, DivSeason)
```

Tukey Tests

```{r}
#Restructured tukey dataset
BeeTukey <- read.csv("BeeTukey.csv")

#Tukey Test
anova1 <- aov(Abundance~Season, data = AllBees)
tukey1 <- TukeyHSD(anova1, conf.level=.95)
tukey1.cld <-multcompLetters4(anova1,tukey1)
summary(anova1)
print(tukey1)
print(tukey1.cld)

report(anova1)

anova2 <- aov(Richness~Season, data = AllBees)
tukey2 <- TukeyHSD(anova2, conf.level=.95)
tukey2.cld <-multcompLetters4(anova2,tukey2)
summary(anova2)
print(tukey2)
print(tukey2.cld)

anova3 <- aov(Shannon_Diversity~Season, data = AllBees)
tukey3 <- TukeyHSD(anova3, conf.level=.95)
tukey3.cld <-multcompLetters4(anova3,tukey3)
summary(anova3)
print(tukey3)
print(tukey3.cld)

#Add tukey test results to dataset
Abundance.cld <- as.data.frame.list(tukey1.cld$'Season')
BeeTukey$Abundance_Tukey <- Abundance.cld$Letters 

Richness.cld <- as.data.frame.list(tukey2.cld$'Season')
BeeTukey$Richness_Tukey <- Richness.cld$Letters 

Diversity.cld <- as.data.frame.list(tukey3.cld$'Season')
BeeTukey$Diversity_Tukey <- Diversity.cld$Letters 

#Save restructured dataset
write.csv(BeeTukey, "BeeTukey.csv")

#Plot
str(BeeTukey)
BeeTukey$Season <- as.factor(BeeTukey$Season)

#Abundance
ggplot(BeeTukey, aes(Season, Abundance_mean, fill =Abundance_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Set2")+
  labs(x = "Season", y = "Average Bee Abundance")+ 
  theme_bw()+
  scale_x_discrete(breaks=c("1PreBloom","2PeakBloom","3DeclineBloom"),
        labels=c("PreBloom", "PeakBloom", "DeclineBloom"))+
  geom_errorbar(aes(ymin=Abundance_mean-Abundance_SE, ymax=Abundance_mean+Abundance_SE), width = 0.2, color = "gray30")+ 
  geom_text(aes(label =Abundance_Tukey), nudge_y = 20, size = 5)+
  ggtitle("Average Bee Abundance by Season")

#Richness
ggplot(BeeTukey, aes(Season, Richness_mean, fill =Richness_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Spectral")+
  labs(x = "Season", y = "Average Bee Richness")+ 
  theme_bw()+
  scale_x_discrete(breaks=c("1PreBloom","2PeakBloom","3DeclineBloom"),
        labels=c("PreBloom", "PeakBloom", "DeclineBloom"))+
  geom_errorbar(aes(ymin=Richness_mean-Richness_SE, ymax=Richness_mean+Richness_SE), width = 0.2, color = "gray30")+ 
  geom_text(aes(label =Richness_Tukey), nudge_y = 1.5, size = 5)+
  ggtitle("Average Bee Species Richness by Season")
  
#Diversity
ggplot(BeeTukey, aes(Season, Shannon_Diversity_mean, fill =Diversity_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Accent")+
  labs(x = "Season", y = "Average Bee Shannon Diversity")+ 
  theme_bw()+
  scale_x_discrete(breaks=c("1PreBloom","2PeakBloom","3DeclineBloom"),
        labels=c("PreBloom", "PeakBloom", "DeclineBloom"))+
  geom_errorbar(aes(ymin=Shannon_Diversity_mean-Shannon_Diversity_SE, ymax=Shannon_Diversity_mean+Shannon_Diversity_SE), width = 0.2, color = "gray30")+
  geom_text(aes(label =Diversity_Tukey), nudge_y = 0.3, size = 5)+
  ggtitle("Average Bee Shannon Diversity by Season")
```

-By Site

```{r}
#Tukey Test
anova4 <- aov(Abundance~Site, data = AllBees)
tukey4 <- TukeyHSD(anova4, conf.level=.95)
tukey4.cld <-multcompLetters4(anova4,tukey4)
print(tukey4)
print(tukey4.cld)

anova5 <- aov(Richness~Site, data = AllBees)
tukey5 <- TukeyHSD(anova5, conf.level=.95)
tukey5.cld <-multcompLetters4(anova5,tukey5)
print(tukey5)
print(tukey5.cld)

anova6 <- aov(Shannon_Diversity~Site, data = AllBees)
tukey6 <- TukeyHSD(anova6, conf.level=.95)
tukey6.cld <-multcompLetters4(anova6,tukey6)
print(tukey6)
print(tukey6.cld)

#Restructured dataset for error bars
BeeSiteTukey <- read.csv("BeeSiteTukey.csv")

#Add tukey test results to dataset
Abundance.cld4 <- as.data.frame.list(tukey4.cld$'Site')
BeeSiteTukey$Abundance_Tukey <- Abundance.cld4$Letters 

Richness.cld5 <- as.data.frame.list(tukey5.cld$'Site')
BeeSiteTukey$Richness_Tukey <- Richness.cld5$Letters 

Diversity.cld6 <- as.data.frame.list(tukey6.cld$'Site')
BeeSiteTukey$Diversity_Tukey <- Diversity.cld6$Letters 

#Save restructured dataset
write.csv(BeeSiteTukey, "BeeSiteTukey.csv")

BeeSiteTukey <- read.csv("BeeSiteTukey.csv")

#Plot
str(BeeSiteTukey)
BeeSiteTukey$Site <- as.factor(BeeSiteTukey$Site)

#Abundance
ggplot(BeeSiteTukey, aes(Site, BeeSiteTukey$Abun_mean, fill = Abundance_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Set2")+
  labs(x = "Site", y = "Average Bee Abundance")+ 
  theme_bw()+
  geom_errorbar(ymin=BeeSiteTukey$Abun_mean-BeeSiteTukey$Abundance_SE, ymax =BeeSiteTukey$Abun_mean+BeeSiteTukey$Abundance_SE, width = 0.2)+ 
  geom_text(aes(label =Abundance_Tukey), nudge_y = 60, size = 5)+
  scale_y_continuous(limits=c(0,150))+
  ggtitle("Average Bee Abundance by Site")

#Richness
ggplot(BeeSiteTukey, aes(Site, Rich_mean, fill =Richness_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Paired")+
  labs(x = "Site", y = "Average Bee Richness")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Rich_mean-Richness_SE, ymax=Rich_mean+Richness_SE), width = 0.2, color = "gray30")+ 
  geom_text(aes(label =Richness_Tukey), nudge_y = 1.5, size = 5)+
  ggtitle("Average Bee Species Richness by Site")
  
#Diversity
ggplot(BeeSiteTukey, aes(Site, Shannon_mean, fill =Diversity_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Accent")+
  labs(x = "Site", y = "Average Bee Shannon Diversity")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Shannon_mean-Shannon_Diversity_SE, ymax=Shannon_mean+Shannon_Diversity_SE), width = 0.2, color = "gray30")+
  geom_text(aes(label =Diversity_Tukey), nudge_y = 0.1, size = 5)+
  ggtitle("Average Bee Shannon Diversity by Site")
```

Floral & Bee Combined Tukeys

```{r}
CombinedSiteTukey <- read.csv("CombinedSiteTukey.csv")
```

-By Season

```{r}
#Abundance
ggplot(CombinedSiteTukey, aes(Site, Abundance_mean, fill = Type))+
  geom_bar(stat= "identity", width =0.8, position=position_dodge())+
  scale_fill_brewer(palette = "Paired")+
  scale_color_brewer(palette = "Paired")+
  labs(x = "Site", y = "Abundance")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Abundance_mean-Abundance_SD, ymax=Abundance_mean+Abundance_SD), width = 0.5, position = position_dodge())+ 
  geom_text(aes(label =Abundance_Tukey, color = Type), position = position_dodge(0.90), size = 6, vjust=-2)+
  ggtitle("Average Abundance by Site")

#Richness
ggplot(CombinedSiteTukey, aes(Site, Richness_mean, fill = Richness_Tukey))+
  geom_bar(stat= "identity", width =0.8, position=position_dodge())+
  scale_fill_brewer(palette = "Paired")+
  scale_color_brewer(palette = "Paired")+
  labs(x = "Site", y = "Richness")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Richness_mean-Richness_SD, ymax=Richness_mean+Richness_SD), width = 0.5, position = position_dodge())+ 
  geom_text(aes(label =Richness_Tukey, color = Type), position = position_dodge(0.90), size = 6, vjust=-2)+
  ggtitle("Average Richness by Site")
#b isnt showing up
  
#Diversity
ggplot(CombinedSiteTukey, aes(Site, Shannon_Diversity, fill =Diversity_Tukey))+
  geom_bar(stat= "identity", width =0.8)+
  scale_fill_brewer(palette = "Paired")+
  labs(x = "Site", y = "Average Bee Shannon Diversity")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Shannon_Diversity-Shannon_Diversity_SD, ymax=Shannon_Diversity+Shannon_Diversity_SD), width = 0.2, color = "gray30")+
  geom_text(aes(label =Diversity_Tukey, color = Diversity_Tukey), nudge_y = 0.1, size = 5)+
  ggtitle("Average Bee Shannon Diversity by Site")
```

-By Site

```{r}
#Abundance
ggplot(BeeSiteTukey, aes(Site, Abundance_mean, fill =Abundance_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Set2")+
  labs(x = "Site", y = "Average Bee Abundance")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Abundance_mean-Abundance_SD, ymax=Abundance_mean+Abundance_SD), width = 0.5, color = "gray30")+ 
  geom_text(aes(label =Abundance_Tukey), nudge_y = 60, size = 5)+
  ggtitle("Average Bee Abundance by Site")

#Richness
ggplot(BeeSiteTukey, aes(Site, Richness_mean, fill =Richness_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Paired")+
  labs(x = "Site", y = "Average Bee Richness")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Richness_mean-Richness_SD, ymax=Richness_mean+Richness_SD), width = 0.2, color = "gray30")+ 
  geom_text(aes(label =Richness_Tukey), nudge_y = 1.5, size = 5)+
  ggtitle("Average Bee Species Richness by Site")
  
#Diversity
ggplot(BeeSiteTukey, aes(Site, Shannon_Diversity_mean, fill =Diversity_Tukey))+
  geom_bar(stat= "identity", width =0.8, show.legend = FALSE)+
  scale_fill_brewer(palette = "Accent")+
  labs(x = "Site", y = "Average Bee Shannon Diversity")+ 
  theme_bw()+
  geom_errorbar(aes(ymin=Shannon_Diversity_mean-Shannon_Diversity_SD, ymax=Shannon_Diversity_mean+Shannon_Diversity_SD), width = 0.2, color = "gray30")+
  geom_text(aes(label =Diversity_Tukey), nudge_y = 0.1, size = 5)+
  ggtitle("Average Bee Shannon Diversity by Site")
```

Least Square Means for sites

BEES

```{r}
#Abundance
Amodel <- lm(Abundance ~ Site, data = AllBees)
anova(Amodel)
#no sig diff between sites
A <- emmeans(Amodel, ~ Site)
write.csv(A, "A.csv")

pairs(A)

pwpm(A)

#Richness 
Rmodel <- lm(Richness ~ Site, data = AllBees)
anova(Rmodel)
#sig diff between sites
R <- emmeans(Rmodel, ~ Site)
write.csv(R, "R.csv")

pairs(R)
# Kamiak - TurnbullSpangle           7.333 1.61 20   4.568  0.0056
# Smoot - TurnbullPlaza              3.667 1.61 20   2.284  0.4386

pwpm(R)

#Diversity
Dmodel <- lm(Shannon_Diversity ~ Site, data = AllBees)
anova(Dmodel)
#sig diff between sites
D <- emmeans(Dmodel, ~ Site)
write.csv(D, "D.csv")

pairs(D)

# Kamiak - TurnbullPlaza           0.99975 0.291 20   3.436  0.0617
# Kamiak - TurnbullSpangle         1.15031 0.291 20   3.954  0.0212
```

PLANTS

```{r}
#Abundance
FAmodel <- lm(Abundance ~ Site, data = AllFloral)
anova(FAmodel)
#no sig diff between sites
FA <- emmeans(FAmodel, ~ Site)

write.csv(t, "t.csv")

pairs(FA)

#Richness 
FRmodel <- lm(Richness ~ Site, data = AllFloral)
anova(FRmodel)
#no sig diff between sites
FR <- emmeans(FRmodel, ~ Site)

write.csv(FR, "FR.csv")

pairs(FR)


#Diversity
FDmodel <- lm(Shannon_Diversity ~ Site, data = AllFloral)
anova(FDmodel)
#no sig diff between sites
FD <- emmeans(FDmodel, ~ Site)

pairs(FD)

write.csv(FD, "FD.csv")
```

Relationship with Canola

```{r}
Canola <- read.csv("CanolaPercentages.csv")
SiteFloral <- read.csv("BySiteFloral.csv")
```

Distance

```{r}
cor.test(Canola$DistancetoCanola, SiteFloral$Abundance)

ggplot() + 
  geom_point(mapping = aes(y = SiteFloral$Abundance, x= Canola$DistancetoCanola, color = Canola$Site)) + 
  geom_smooth(mapping = aes(y = SiteFloral$Abundance, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "Floral Abundance", x = "Distance to Canola", colour = "Site") +
  ggtitle("Floral Abundance and Distance to Canola")

cor.test(Canola$DistancetoCanola, SiteFloral$Richness)

ggplot() + 
  geom_point(mapping = aes(y = SiteFloral$Richness, x= Canola$DistancetoCanola, color = Canola$Site)) + 
  geom_smooth(mapping = aes(y = SiteFloral$Richness, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "Floral Richness", x = "Distance to Canola", colour = "Site") +
  ggtitle("Floral Richness and Distance to Canola")

cor.test(Canola$DistancetoCanola, SiteFloral$Shannon_Diversity)

ggplot() + 
  geom_point(mapping = aes(y = SiteFloral$Shannon_Diversity, x= Canola$DistancetoCanola, color = Canola$Site)) + 
  geom_smooth(mapping = aes(y = SiteFloral$Shannon_Diversity, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "Floral Shannon Diversity", x = "Distance to Canola", colour = "Site") +
  ggtitle("Floral Diversity and Distance to Canola")
```

Corr between bee div and plant div

```{r}
AllFloral <- read.csv("AllFloral.csv")
AllBees <- read.csv("BeesbySiteSeason.csv")


AbunAbun <- cor.test(AllFloral$Abundance, AllBees$Abundance)
#p-value = 0.7365
summary(lm(AllBees$Abundance~AllFloral$Abundance))
summary(lmer(AllBees$Abundance~AllFloral$Abundance + (1|AllFloral$Site)))

AbunRich <- cor.test(AllFloral$Abundance, AllBees$Richness)
#p-value = 0.2721
summary(lm(AllBees$Richness~AllFloral$Abundance))
summary(lmer(AllBees$Richness~AllFloral$Abundance + (1|AllFloral$Site)))

AbunDiv <- cor.test(AllFloral$Abundance, AllBees$Shannon_Diversity)
#p-value = 0.5245
summary(lm(AllBees$Shannon_Diversity~AllFloral$Abundance))
summary(lmer(AllBees$Shannon_Diversity~AllFloral$Abundance + (1|AllFloral$Site)))

report(AbunDiv)
#floral abundance does not increase bee abundance, richness, or diversity


RichAbun <- cor.test(AllFloral$Richness, AllBees$Abundance)
#p-value = 0.1176
summary(lm(AllBees$Abundance~AllFloral$Richness))
summary(lmer(AllBees$Abundance~AllFloral$Richness + (1|AllFloral$Site)))

RichRich <- cor.test(AllFloral$Richness, AllBees$Richness)
#p-value = 0.4744
summary(lm(AllBees$Richness~AllFloral$Richness))
summary(lmer(AllBees$Richness~AllFloral$Richness + (1|AllFloral$Site)))

RichDiv <- cor.test(AllFloral$Richness, AllBees$Shannon_Diversity)
#p-value = 0.5027
summary(lm(AllBees$Shannon_Diversity~AllFloral$Richness))
summary(lmer(AllBees$Shannon_Diversity~AllFloral$Richness + (1|AllFloral$Site)))

#floral richness does not increase bee abundance, richness, or diversity

cor.test(AllFloral$Shannon_Diversity, AllBees$Abundance)
#p-value = 0.4771
summary(lm(AllBees$Abundance~AllFloral$Shannon_Diversity))
summary(lmer(AllBees$Abundance~AllFloral$Shannon_Diversity + (1|AllFloral$Site)))

cor.test(AllFloral$Shannon_Diversity, AllBees$Richness)
#p-value = 0.5315
summary(lm(AllBees$Richness~AllFloral$Shannon_Diversity))
summary(lmer(AllBees$Richness~AllFloral$Shannon_Diversity + (1|AllFloral$Site)))

cor.test(AllFloral$Shannon_Diversity, AllBees$Shannon_Diversity)
#p-value = 0.7075
summary(lm(AllBees$Shannon_Diversity~AllFloral$Shannon_Diversity))
summary(lmer(AllBees$Shannon_Diversity~AllFloral$Shannon_Diversity + (1|AllFloral$Site)))
#floral div does not increase bee abundance, richness, or diversity

#No relationship between plant div and bee div
```
