---
title: "BeeStability"
author: "Olivia Shaffer"
date: "1/7/2023"
output: html_document
---
https://cran.r-project.org/web/packages/codyn/vignettes/Community_Stability_Metrics.html

```{r}
install.packages("knitr")
install.packages("codyn")
install.packages("tidyverse")
install.packages("vegan")
install.packages("ggplot2")
install.packages("ggThemeAssist")
```

```{r}
library(codyn)
library(knitr)
library(tidyverse)
library(vegan)
library(ggplot2)
library(ggThemeAssist)
```

Looking at given data
```{r}
data(knz_001d)
#Make a datasheet that's structured like this

#Abundance of each species at each transect
##Try site first
```

Making sure the function works
```{r}
KNZ_stability <- community_stability(knz_001d, 
                                   time.var = "year", 
                                   abundance.var = "abundance", 
                                   replicate.var = "subplot")
View(KNZ_stability)
```

Making Abundance dataset
```{r}
bees <- read.csv("Bee_Samples_2022.csv")
```

Create Frequency tables
```{r}
#SiteSeason column
bees$SiteSeason <- paste(bees$Site, bees$Period, sep="_")

Freq <- table(bees$SiteSeason, bees$Genus)
Freq2 <- as.data.frame(Freq)

Freq2 %>% rename(SiteSeason = Var1, Genus = Var2, Abundance = Freq)
View(Freq2)

write.csv(Freq2, "AbundanceAll.csv")
```


Load in edited Abundance dataset
```{r}
Abundance <- read.csv("AbundanceReformat.csv")
Abundance$SamplingPeriod <- as.numeric(Abundance$Sampling.Period)
```


Stability
```{r}
#Abundance by species or Richness
stability <- community_stability(Abundance,
                                   time.var = "SamplingPeriod", 
                                   abundance.var = "Abundance", 
                                   replicate.var = "Site")

View(stability)

```

Save 
```{r}
write.csv(stability, "StabilitybySite.csv")
stability <- read.csv("StabilitybySite.csv")
```
-----------------------------------------------------------------------------
Test values

```{r}
#Load in fake data
test <- read.csv("TestData.csv")
test$SamplingPeriod <- as.numeric(test$SamplingPeriod)
test$Lowest_Stab <- as.numeric(test$Lowest_Stab)
```

```{r}
#Abundance by species or Richness
test_stability <- community_stability(test,
                                   time.var = "SamplingPeriod", 
                                   abundance.var = "Lowest_Stab",
                                   replicate.var = "Site")
View(test_stability)
```
-----------------------------------------------------------------------------
Correlation between canola and stability
```{r}
#read in landscape data
Canola <- read.csv("CanolaPercentages.csv")

#Combined landscape & Canola data for linear regressions

All <- read.csv("BeesStabilityLandscape.csv")
```

```{r}
#Same values whether you're using canola area or %

#canola % within 2km
#cor.test(Canola$X2kmCan., stability$stability)
#p-value = 0.06846
# cor -0.5969427  

#linear regression
summary(lm(All, stability ~ X2kmCan. + (1|Site)))


ggplot() + 
  geom_point(mapping = aes(y = stability$stability, x= Canola$X2kmCan., color = stability$Site)) + 
  geom_smooth(mapping = aes(y = stability$stability, x = Canola$X2kmCan.), method = lm) +
  labs(y = "Stability", x = "2km Canola %", colour = "Site") +
  ggtitle("Bee Variation in Abundance and 2km Canola %")


#canola % within 1km
#cor.test(Canola$X1kmCan., Canola$stability)
#p-value = 0.2754
# cor -0.3824552

#linear regression
summary(lm(stability ~ X1kmCan. + (1|Site)))

ggplot() + 
  geom_point(mapping = aes(y = Canola$stability, x = Canola$X1kmCan., color = Canola$Site)) + 
  labs(y = "Community Stability", x = "Canola % in 1km", colour = "Site") +
  annotate(geom="text", x=50, y=0.6, label="P= 0.2754") +
  ggtitle("Bee Variation in Abundance and 1km Canola %")

#canola % within 500m
#cor.test(Canola$X500mCan., Canola$stability)
#p-value = 0.9443
#cor -0.02549981 

#linear regression
summary(lm(stability ~ X500mCan. + (1|Site)))

ggplot() + 
  geom_point(mapping = aes(y = Canola$stability, x = Canola$X500mCan., color = Canola$Site)) + 
  labs(y = "Community Stability", x = "Canola % in 500m", colour = "Site") +
  annotate(geom="text", x=30, y=1.6, label="P= 0.9443") +
  ggtitle("Relationship between Bee Stability and Canola in 500m")

##All distances on one boxplot
ggplot() + 
  geom_boxplot()
  labs(y = "Stability", x = "Canola %") +
  ggtitle("Canola % up to 2km Does Not Effect Bee Variation in Abundance")

#canola distance
#cor.test(Canola$DistancetoCanola, stability$stability)
#p-value = 0.5226
#cor -0.230039 

#linear regression
summary(lm(stability ~ DistancetoCanola + (1|Site)))

ggplot() + 
  geom_point(mapping = aes(y = stability$stability, x = Canola$DistancetoCanola, color = stability$Site)) + 
  geom_smooth(mapping = aes(y = stability$stability, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "Stability", x = "Distance to Canola (m)", colour = "Site") +
  ggtitle("Bee Variation in Abundance and Distance to Canola")

```


T-tests & Boxplots
-Above 20%, Below 20%
-Within flight distance (>500km), 
```{r}
t.test(stability$stability ~ Canola$Canolain2km)
#P=0.09351
#Above 20% mean = 1.216596
#Below 20% mean = 2.105196 

ggplot() +
  geom_boxplot(aes(x=reorder(Canola$Canolain2km,-stability$stability), y=stability$stability, color=Canola$Canolain2km))+
  labs(y = "Stability", x = "Percent of Canola in 2km")+
  theme(legend.position = "none") +
  ggtitle("Bee Variation in Abundance and 2km Canola %")

t.test(stability$stability ~Canola$FlightDistance)
#P=0.9606
#Above mean= 1.76284
#within mean= 1.73013

ggplot() +
  geom_boxplot(aes(x=reorder(Canola$FlightDistance,stability$stability),y=stability$stability, color=Canola$FlightDistance))+
  labs(y = "Stability", x = "Relation of Canola to Bee Flight Distance(500m)")+
  theme(legend.position = "none") +
  ggtitle("Bee Variation in Abundance and Canola within Flight Distance")


```


##Canola and PDI
```{r}
#read in PDI
PDI <- read.csv("PDIbySite.csv")
```

```{r}
#canola % within 2km
cor.test(Canola$X2kmCan., PDI$PDI)
#p-value = 0.1945
# cor 0.4476353  

ggplot() + 
  geom_point(mapping = aes(y = PDI$PDI, x= Canola$X2kmCan., color = stability$Site)) + 
  geom_smooth(mapping = aes(y = PDI$PDI, x = Canola$X2kmCan.), method = lm) +
  labs(y = "PDI", x = "2km Canola %", colour = "Site") +
  ggtitle("Bee PDI and 2km Canola %")

#canola % within 1km
cor.test(Canola$X1kmCan., Canola$PDI)
#p-value = 0.2429
#cor 0.407109 

ggplot() + 
  geom_point(mapping = aes(y = Canola$PDI, x = Canola$X1kmCan., color = Canola$Site)) + 
  labs(y = "PDI", x = "Canola % in 1km", colour = "Site") +
  annotate(geom="text", x=50, y=0.6, label="P= 0.2429") +
  ggtitle("Relationship between PDI and Canola in 1km")


#canola % within 500m
cor.test(Canola$X500mCan., Canola$PDI)
#p-value = 0.6055
# cor 0.1867166 

ggplot() + 
  geom_point(mapping = aes(y = Canola$PDI, x = Canola$X500mCan., color = Canola$Site)) + 
  labs(y = "PDI", x = "Canola % in 500m", colour = "Site") +
  annotate(geom="text", x=30, y=0.6, label="P= 0.6055") +
  ggtitle("Relationship between PDI and Canola in 500m")

##Bar graph showing all distances


#canola distance
cor.test(Canola$DistancetoCanola, Canola$PDI)
#p-value = 0.531
#cor 0.2255469 


#PDI
ggplot() + 
  geom_point(mapping = aes(y = PDI$PDI, x = Canola$DistancetoCanola, color = stability$Site)) + 
  geom_smooth(mapping = aes(y = PDI$PDI, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "PDI", x = "Distance to Canola (m)", colour = "Site") +
  ggtitle("Bee PDI and Distance to Canola")


#All Together
ggplot() + 
  geom_point(mapping = aes(y = PDI$PDI, x = Canola$DistancetoCanola)) + 
  geom_smooth(mapping = aes(y = PDI$PDI, x = Canola$DistancetoCanola, color = 'PDI'), method = lm) +
  geom_point(mapping = aes(y = stability$stability, x = Canola$DistancetoCanola)) + 
  geom_smooth(mapping = aes(y = stability$stability, x = Canola$DistancetoCanola, color = 'Variation in
Abundance'), method = lm) +
  theme_bw()+
  labs(y = "Bee Stability", x = "Distance to Canola (m)", colour = "Site") +
  ggtitle("Bee Stability and Distance to Canola")

ggplot() + 
  geom_point(mapping = aes(y = PDI$PDI, x = Canola$X2kmCan.)) + 
  geom_smooth(mapping = aes(y = PDI$PDI, x = Canola$X2kmCan., color = 'PDI'), method = lm) +
  geom_point(mapping = aes(y = stability$stability, x = Canola$X2kmCan.)) + 
  geom_smooth(mapping = aes(y = stability$stability, x = Canola$X2kmCan., color = 'Variation in 
Abundance'), method = lm) +
  theme_bw()+
  labs(y = "Bee Stability", x = "2km Canola %", colour = "Site") +
  ggtitle("Bee Stability and Canola %")
```


T-tests and boxplots
```{r}
t.test(PDI$PDI ~ Canola$Canolain2km)
#P=0.1499
#Above 20% mean = 0.9308113 
#Below 20% mean = 0.8122594 

ggplot() +
  geom_boxplot(aes(x=reorder(Canola$Canolain2km,PDI$PDI), y=PDI$PDI, color=Canola$Canolain2km))+
  labs(y = "PDI", x = "Percent of Canola in 2km")+
  theme(legend.position = "none") +
  ggtitle("Bee PDI and 2km Canola %")

t.test(PDI$PDI ~Canola$FlightDistance)
#P=0.9606
#Above mean= 1.76284
#within mean= 1.73013

ggplot() +
  geom_boxplot(aes(x=reorder(Canola$FlightDistance,-PDI$PDI), PDI$PDI, color=Canola$FlightDistance))+
  labs(y = "Stability", x = "Relation of Canola to Bee Flight Distance(500m)")+
  theme(legend.position = "none")+
  ggtitle("Bee PDI and Canola within Flight Distance")

```

##Correlation between PDI and community stability
```{r}
cor.test(stability$stability, PDI$PDI)
#p-value = 0.1658
#cor -0.4745098 
```

##Stability v Other Landscape Coverage
```{r}
land <- read.csv("LandcoverPercentages2022.csv")
```

Stability
```{r}
cor.test(land$Non.flowering.ag, stability$stability)
#p-value = 0.2229
cor.test(land$Barren, stability$stability)
#p-value = 0.8888
cor.test(land$Development, stability$stability)
#p-value = 0.3113
cor.test(land$Tree.Fruit.Crops, stability$stability)
#p-value = 0.3878
cor.test(land$Forest.and.Tree.Crops, stability$stability)
#p-value = 0.9185
cor.test(land$Wetland.and.Open.Water, stability$stability)
#p-value = 0.8858
cor.test(land$Pasture, stability$stability)
#p-value = 0.4336
cor.test(land$Other.Flowering.ag, stability$stability)
#p-value = 0.3859

cor.test(Canola$Canola2km, Canola$stability)
#p-value = 0.06846

ggplot() + 
  geom_point(mapping = aes(y = stability$stability, x = land$Non.flowering.ag, color = 'Non Flowering Ag')) + 
  geom_point(mapping = aes(y = stability$stability, x = land$Barren, color = 'Barren')) + 
  geom_point(mapping = aes(y = stability$stability, x = land$Development, color = 'Development')) + 
  geom_point(mapping = aes(y = stability$stability, x = land$Tree.Fruit.Crops, color = 'Tree Fruit')) + 
  geom_point(mapping = aes(y = stability$stability, x = land$Forest.and.Tree.Crops, color = 'Forest')) + 
  geom_point(mapping = aes(y = stability$stability, x = land$Wetland.and.Open.Water, color = 'Wetland and Water')) + 
  geom_point(mapping = aes(y = stability$stability, x = land$Pasture, color = 'Pasture')) + 
  geom_point(mapping = aes(y = stability$stability, x = land$Other.Flowering.ag, color = 'Other Flowering ag')) + 
  geom_smooth(mapping = aes(y = Canola$stability, x = land$Manually.CalculatedCanola, color = 'Canola'), method = lm) +
  labs(y = "Stability", x = "Percentage of Landscape Type", colour = "Landscape Type") +
  ggtitle("Bee Stability and Background Landscape Types")

```
PDI
```{r}
cor.test(land$Non.flowering.ag, PDI$PDI)
#p-value = 0.9952
cor.test(land$Barren, PDI$PDI)
#p-value = 0.9625
cor.test(land$Development, PDI$PDI)
#p-value = 0.5806
cor.test(land$Tree.Fruit.Crops, PDI$PDI)
#p-value = 0.2849
cor.test(land$Forest.and.Tree.Crops, PDI$PDI)
#p-value = 0.355
cor.test(land$Wetland.and.Open.Water, PDI$PDI)
#p-value = 0.2577
cor.test(land$Pasture, PDI$PDI)
#p-value = 0.7966
cor.test(land$Other.Flowering.ag, PDI$PDI)
#p-value = 0.364

cor.test(Canola$Canola2km, Canola$PDI)
#p-value = 0.1945

ggplot() + 
  geom_point(mapping = aes(y = PDI$PDI, x = land$Non.flowering.ag, color = 'Non Flowering Ag')) + 
  geom_point(mapping = aes(y = PDI$PDI, x = land$Barren, color = 'Barren')) + 
  geom_point(mapping = aes(y = PDI$PDI, x = land$Development, color = 'Development')) + 
  geom_point(mapping = aes(y = PDI$PDI, x = land$Tree.Fruit.Crops, color = 'Tree Fruit')) + 
  geom_point(mapping = aes(y = PDI$PDI, x = land$Forest.and.Tree.Crops, color = 'Forest')) + 
  geom_point(mapping = aes(y = PDI$PDI, x = land$Wetland.and.Open.Water, color = 'Wetland and Water')) + 
  geom_point(mapping = aes(y = PDI$PDI, x = land$Pasture, color = 'Pasture')) + 
  geom_point(mapping = aes(y = PDI$PDI, x = land$Other.Flowering.ag, color = 'Other Flowering ag')) +
  geom_point(mapping = aes(y = Canola$PDI, x = land$Manually.CalculatedCanola, color = 'Canola')) +
  labs(y = "PDI", x = "Percentage of Landscape Type", colour = "Landscape Type") +
  ggtitle("PDI and Background Landscape Types")
```

Correlation Between Stability and Plant Diversity
```{r}
Floral <- read.csv("BySiteFloral.csv")
```

```{r}
#Abundance
cor.test(Floral$Abundance, stability$stability)
#p-value = 0.486
#cor=0.2500465

#Richness
cor.test(Floral$Richness, stability$stability)
#p-value = 0.6033
#cor= -0.1878432 

#Diversity
cor.test(Floral$Shannon_Diversity, stability$stability)
#p-value = 0.5881
#cor= -0.1956252 


#Abundance
cor.test(Floral$Abundance, PDI$PDI)
#p-value = 0.3457
#cor = 0.3339429 

#Richness
cor.test(Floral$Richness, PDI$PDI)
#p-value = 0.1805
#cor=0.4604281 

#Diversity
cor.test(Floral$Shannon_Diversity, PDI$PDI)
#p-value = 0.02792
#cor=0.687828 
```

Correlation between bee stability & plany stability
```{r}
cor.test(BeeStability$PDI, Plants$PDI)
#t = 3.4927, df = 8, p-value = 0.008166
#cor = 0.7771344 
cor.test(BeeStability$Stability, Plants$Stability)
#t = -0.95952, df = 8, p-value = 0.3654
#cor = -0.3212587
cor.test(BeeStability$PDI, Plants$Stability)
#t = 0.3734, df = 8, p-value = 0.7185
# cor = 0.1308822 
cor.test(BeeStability$Stability, Plants$PDI)
#t = 0.29205, df = 8, p-value = 0.7777
#cor = 0.1027101

ggplot() +
   geom_point(mapping = aes(y = BeeStability$PDI, x = Plants$PDI, color = Plants$Site)) +
  labs(y = "Bee PDI", x = "Floral PDI") +
  ggtitle("Bee Community PDI and Floral PDI")
```


Stability 

```{r}
#Abundance
ggplot() +
  geom_point(mapping = aes(y = stability$stability, x = Floral$Abundance, color = 'red')) +
  labs(y = "Bee Stability", x = "Floral Abundance") +
  theme(legend.position = "none") +
  annotate(geom="text", x=700, y= 2.5, label="P=0.486") +
  ggtitle("Relationship between Bee Community Stability and Floral Abundance")

#Diversity
ggplot() +
   geom_point(mapping = aes(y = stability$stability, x = Floral$Richness, color = 'green')) +
  theme(legend.position = "none") +
  annotate(geom="text", x=20, y= 1, label="P=0.6033") +
  labs(y = "Bee Stability", x = "Floral Richness") +
  ggtitle("Relationship between Bee Community Stability and Floral Richness")

#Diversity
ggplot()+
geom_point(mapping = aes(y = stability$stability, x = Floral$Shannon_Diversity, color = 'blue')) +  theme(legend.position = "none") +
  annotate(geom="text", x=2.75, y= 2.5, label="P= 0.5881") +
  labs(y = "Bee Stability", x = "Floral Shannon Diversity") +
  ggtitle("Relationship between Bee Community Stability and Floral Diversity")
```

PDI
```{r}
#Abundance
ggplot() +
  geom_point(mapping = aes(y = PDI$PDI, x = Floral$Abundance, color = 'red')) +
  labs(y = "PDI", x = "Floral Abundance") +
  theme(legend.position = "none") +
  annotate(geom="text", x=700, y= 1.00, label="0.3457") +
  ggtitle("Relationship between Bee PDI and Floral Abundance")

#Diversity
ggplot() +
  geom_point(mapping = aes(y = PDI$PDI, x = Floral$Richness, color = 'green')) +
  theme(legend.position = "none") +
  annotate(geom="text", x=22, y= 1, label="0.1805") +
  labs(y = "PDI", x = "Floral Richness") +
  ggtitle("Relationship between Bee PDI and Floral Richness")

#Diversity
ggplot()+
geom_smooth(mapping = aes(y = PDI$PDI, x = Floral$Shannon_Diversity, color = 'blue'), method = lm) +
  geom_point(mapping = aes(y = PDI$PDI, x = Floral$Shannon_Diversity, color = 'blue'))+
theme(legend.position = "none") +
  annotate(geom="text", x=2.75, y= 0.6, label="P= 0.02792") +
  labs(y = "PDI", x = "Floral Shannon Diversity") +
  ggtitle("Relationship between Bee PDI and Floral Diversity")
```

#Bee Diversity Metrics & Canola
```{r}
BeeDiv <- read.csv("BeesbySite.csv")
```

Canola %

-Pearson's correlation
```{r}
#Richness
cor.test(Canola$X2kmCan., BeeDiv$Richness)
#p-value = 0.5949
# cor -0.1921248  

ggplot() + 
  geom_point(mapping = aes(y = BeeDiv$Richness, x= Canola$X2kmCan., color = BeeDiv$Site)) + 
  geom_smooth(mapping = aes(y = BeeDiv$Richness, x = Canola$X2kmCan.), method = lm) +
  labs(y = "Bee Richness", x = "2km Canola %", colour = "Site")+
  ggtitle("Bee Richness and Canola % within 2km")

cor.test(Canola$X1kmCan., BeeDiv$Richness)
#p-value = 0.5377
#cor -0.2219471 

cor.test(Canola$X500mCan., BeeDiv$Richness)
#p-value = 0.5137
#cor -0.2348577 


#Diversity
cor.test(Canola$X2kmCan., BeeDiv$Shannon_Diversity)
#p-value = 0.9503
# cor -0.02273696   

ggplot() + 
  geom_point(mapping = aes(y = BeeDiv$Shannon_Diversity, x= Canola$X2kmCan., color = BeeDiv$Site)) + 
  geom_smooth(mapping = aes(y = BeeDiv$Shannon_Diversity, x = Canola$X2kmCan.), method = lm) +
  labs(y = "Shannon Diversity", x = "2km Canola %", colour = "Site") +
  ggtitle("Bee Shannon Diversity and Canola % within 2km")

cor.test(Canola$X1kmCan., BeeDiv$Shannon_Diversity)
#p-value = 0.7638
#cor -0.1092634 

#Abundance
cor.test(Canola$X2kmCan., BeeDiv$Abundance)
#p-value = 0.4351
# cor 0.278939    

ggplot() + 
  geom_point(mapping = aes(y = BeeDiv$Abundance, x= Canola$X2kmCan., color = BeeDiv$Site)) + 
  geom_smooth(mapping = aes(y = BeeDiv$Abundance, x = Canola$X2kmCan.), method = lm) +
  labs(y = "Abundance", x = "2km Canola %", colour = "Site") +
  ggtitle("Bee Abundance and Canola % within 2km")

#Canola percentage x Diversity combined

ggplot() + 
   geom_point(mapping = aes(y = BeeDiv$Shannon_Diversity, x= Canola$X2kmCan.)) + 
  geom_smooth(mapping = aes(y = BeeDiv$Shannon_Diversity, x = Canola$X2kmCan., color = 'Shannon Diversity'), method = lm) +
   geom_point(mapping = aes(y = BeeDiv$Richness, x= Canola$X2kmCan.)) + 
  geom_smooth(mapping = aes(y = BeeDiv$Richness, x = Canola$X2kmCan., color = 'Richness'), method = lm) +
  labs(y = "Bee Diversity Metrics", x = "2km Canola %") +
  ggtitle("Bee Diversity and Canola % within 2km")
```

t.test
```{r}
#Richness
t.test(BeeDiv$Richness ~ Canola$Canolain2km)
#P=0.5489
#Above 20% mean = 10.50000 
#Below 20% mean = 11.66667 

ggplot() +
  geom_boxplot(aes(x=reorder(Canola$Canolain2km, -BeeDiv$Richness), y=BeeDiv$Richness, color=Canola$Canolain2km))+
  annotate(geom="text", x=2.3, y=0.6, label="P=0.5489")+
  labs(y = "Bee Richness", x = "Percent of Canola in 2km")+
  theme(legend.position = "none") +
  ggtitle("Bee Richness and 2km Canola %")

#Diversity
t.test(BeeDiv$Shannon_Diversity~ Canola$Canolain2km)
#P=0.5054
#Above mean= 1.794138    
#within mean= 1.640124 

ggplot() +
  geom_boxplot(aes(x=reorder(Canola$Canolain2km,BeeDiv$Shannon_Diversity), y= BeeDiv$Shannon_Diversity, color=Canola$Canolain2km))+
  annotate(geom="text", x=2.3, y=0.6, label="P= 0.5054")+
  labs(y = "Shannon Diversity", x = "Percent of Canola in 2km")+
  theme(legend.position = "none") +
  ggtitle("Bee Diversity and 2km Canola %")

#Abundance
t.test(BeeDiv$Abundance~ Canola$Canolain2km)
#P=0.851
#Above mean= 135.5    
#within mean= 127.0  

ggplot() +
  geom_boxplot(aes(x=reorder(Canola$Canolain2km,BeeDiv$Abundance), y= BeeDiv$Abundance, color=Canola$Canolain2km))+
  annotate(geom="text", x=2.2, y=0.6, label="P = 0.851")+
  labs(y = "Bee Abundance", x = "Percent of Canola in 2km")+
  theme(legend.position = "none") +
  ggtitle("Bee Abundance and 2km Canola %")

```

Canola Distance
-Pearson's Correlation
```{r}
#Richness
cor.test(Canola$DistancetoCanola, BeeDiv$Richness)
#p-value = 0.836
#cor -0.07541681 

ggplot() + 
  geom_point(mapping = aes(y = BeeDiv$Richness, x = Canola$DistancetoCanola, color = BeeDiv$Site)) + 
  geom_smooth(mapping = aes(y = BeeDiv$Richness, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "Richness", x = "Distance to Canola (m)", colour = "Site")+ 
  ggtitle("Bee Richness and Distance to Canola")

#Diversity
cor.test(Canola$DistancetoCanola, BeeDiv$Shannon_Diversity)
#p-value = 0.9584
#cor 0.019044099 

ggplot() + 
  geom_point(mapping = aes(y = BeeDiv$Shannon_Diversity, x = Canola$DistancetoCanola, color = BeeDiv$Site)) + 
  geom_smooth(mapping = aes(y = BeeDiv$Shannon_Diversity, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "Shannon Diversity", x = "Distance to Canola (m)", colour = "Site") +
  ggtitle("Bee Shannon Diversity and Distance to Canola")

#Abundance
cor.test(Canola$DistancetoCanola, BeeDiv$Abundance)
#p-value = 0.1762
# cor -0.4645298   

ggplot() + 
  geom_point(mapping = aes(y = BeeDiv$Abundance, x= Canola$DistancetoCanola, color = BeeDiv$Site)) + 
  geom_smooth(mapping = aes(y = BeeDiv$Abundance, x = Canola$DistancetoCanola), method = lm) +
  labs(y = "Abundance", x = "Distance to Canola", colour = "Site") +
  ggtitle("Bee Abundance and Distance to Canola")

```
t.test
```{r}
#Richness
t.test(BeeDiv$Richness ~ Canola$FlightDistance)
#P=0.4149
#Above mean = 11.83333  
#Below mean = 10.25000  

ggplot() +
  geom_boxplot(aes(x=reorder(Canola$FlightDistance, BeeDiv$Richness), y=BeeDiv$Richness, color=Canola$FlightDistance))+
  annotate(geom="text", x=2.3, y=0.6, label="P=0.4149")+
  labs(y = "Bee Richness", x = "Relation of Canola to Bee Flight Distance(500m)")+
  theme(legend.position = "none") +
  ggtitle("Bee Richness and Distance to Canola")

#Diversity
t.test(BeeDiv$Shannon_Diversity~ Canola$FlightDistance)
#P=0.5871
#Above mean= 1.756626    
#within mean= 1.619384 

ggplot() +
  geom_boxplot(aes(x=reorder(Canola$FlightDistance ,BeeDiv$Shannon_Diversity), y= BeeDiv$Shannon_Diversity, color=Canola$FlightDistance))+
  annotate(geom="text", x=2.3, y=0.6, label="P= 0.5054")+
  labs(y = "Shannon Diversity", x = "Relation of Canola to Bee Flight Distance(500m)")+
  theme(legend.position = "none") +
  ggtitle("Bee Diversity and Distance to Canola")

#Abundance
t.test(BeeDiv$Abundance~ Canola$FlightDistance)
#P=0.7441
#Above mean=  124.50    
#within mean= 139.25  

ggplot() +
  geom_boxplot(aes(x=reorder(Canola$FlightDistance,BeeDiv$Abundance), y= BeeDiv$Abundance, color=Canola$FlightDistance))+
  annotate(geom="text", x=2.3, y=0.6, label="P= 0.7441")+
  labs(y = "Abundance", x = "Relation of Canola to Bee Flight Distance(500m)")+
  theme(legend.position = "none") +
  ggtitle("Bee Abundance and Distance to Canola")

```


#By site & Season Graphs

Stability Metrics by Site
```{r}
ggplot(stability, aes(x = Site, y=stability, fill = 'skyblue4')) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  theme(legend.position = "none") +
  ggtitle("Bee Community Stability by Field Site")

ggplot(PDI, aes(x = Site, y=PDI, fill = 'skyblue4')) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  theme(legend.position = "none") +
  ggtitle("Bee Dissimilarity by Field Site")
```



